### 2019-11-15 data_extraction_form/instructions.Rmd
convert text file to markdown file. Delete word_document output from knitting markdown file

---
title: "Instructions: data extraction from epidemiology articles"
author: "Katie Drax"
date: "15 November 2019"
output: word_document
---

# Time of writing

- The study designs in **[number?]** articles have been identified by KD and MG. 
- Data extraction has not yet begun for any eligible articles
- The first draft of the data extraction form has been piloted on one observational epidemiology study by MG, KD and RR (Dregan (2018) Arterial stiffness association with chronic inflammatory disorders in the UK Biobank study).

# Procedure

The following procedure will be used for all articles classified as 'observational epidemiology'

1. Two people (from KD, RR, MG and BW) will independently identify the study design(s) (cohort, case-control or cross-sectional) used by each of the 178 articles by answering the first block of questions in the data extraction form. They will also check all supplementary material can be accessed via University of Bristol subscriptions. We will attempt to retrieve any supplementary material we cannot access (e.g., by contacting the authors). Any conflicts will be discussed by the two extractors and resolved by mutual consent if possible. If not resolved after discussion, RR or MM will determine the study designs
1. Articles containing multiple different study designs will be discussed on a case-by-case basis to confirm whether a single reporting checklist is adequate, or if multiple reporting checklists apply to a single article.
1. KD will identify and exclude retracted articles by assessing the notices attached to the online version of the article.
1. The information in Table 1 will be extracted manually by KD or automatically by Endnote.
1. The articles will be randomly selected for data extraction and two people (from MG, KD, BW and RR) will independently complete the data extraction form for each article. Any conflicts will be discussed by the two extractors and resolved by mutual consent if possible. If not resolved after discussion, RR or MM will assign the category. Articles will continue to be selected at random until none remain, or until the authors resources are exhausted.

*Table 1. Methods for extracting article meta-data*

| Item | Definition | Method of extraction |
| --- | --- | --- |
| Article ID(s) of related articles | Article IDs of articles that contain similar analyses. | Examine articles with same first author, last author and UK Biobank application number. |
| Citation |   | Endnote |
| Publication date |   | Endnote |
| Authors |   |  Endnote |
| Journal |   | Endnote  |
| Volume |   | Endnote  |
| Issue |   | Endnote  |
| Open Access |   | (1) Search for article on open access button ([https://openaccessbutton.org/](https://openaccessbutton.org/)) and follow the link on non-university network  (2) If no on open access button check if there are links on Google Scholar |
| Countries of corresponding, first &amp; last author |   | Extract country of author's affiliations |
| UK Biobank Application number | Study's approved UK Biobank data application number (if present) | (1) Search article for all these terms:  'application' , 'approv' , 'number' , 'Resource' (2) If not found examine any strings of numbers in article |
| Author keywords |   | (1) Extract from page 1 (2) If not on page 1 search article for 'keywords' |
| Journal endorsement | Journal's endorsement of STROBE before publication | (1) Extract references to STROBE from the journal's most recent 'instructions/guide for authors' before article's publication date |
| Journal impact factor |   | Retrieve the impact factor of the journal for year of publication from Web of Science's Journal Citation Reports (JCR) |
| Corresponding author email address |   | Endnote; if missing: (1) extract from page 1 (2) If not on page 1 search article for 'corresponding', '@') |
| Conflicts of interest statement |   | (1) Search article material for all these terms: 'competing', 'interest' ,'conflict' , 'disclosure' (2) If not found examine article for similar terms (3) if present extract conflicts of interest statements |
| Classification | Classification ('mendelian randomisation', 'observational epidemiology' or 'other') as assigned by KD, BW and RR in Rayyan | Rayyan |

# Missing items

We will not attempt to retrieve any missing items in the data extraction form, in order to give an accurate assessment of the version of record and to avoid exhausting the extractors resources.

### 2019-11-15 data_extraction_form/instructions.Rmd
add new line after each full stop

@@
 - Data extraction has not yet begun for any eligible articles
-- The first draft of the data extraction form has been piloted on one observational epidemiology study by MG, KD and RR (Dregan (2018) Arterial stiffness association with chronic inflammatory disorders in the UK Biobank study).
+- The Data Extraction form has been piloted on **[number]** observational epidemiology articles by MG, KD and RR. 
+**[X] changes were made so the Data Extraction Form **[will/will not]** be completed again for the piloted articles.
 
@@
 1.
-Two people (from KD, RR, MG and BW) will independently identify the study design(s) (cohort, case-control or cross-sectional) used by each of the 178 articles by answering the first block of questions in the data extraction form.
-They will also check all supplementary material can be accessed via University of Bristol subscriptions.
+Two people (from KD, RR, MG and BW) will independently identify the study design(s) (cohort, case-control or cross-sectional) used by each of the 178 articles by completing the Study Design Form for each article.
+The form also asks extractors to indicate if all supplementary material can be accessed via University of Bristol subscriptions. 
 We will attempt to retrieve any supplementary material we cannot access (e.g., by contacting the authors). 
 Any conflicts will be discussed by the two extractors and resolved by mutual consent if possible. 
-If not resolved after discussion, RR or MM will determine the study designs
+If not resolved after discussion, RR or MM will determine the study designs. 
+The Study Design Form was hosted on Qualtrics but a PDF version can be found here https://osf.io/nd78g/.  
 1.
@@
 1. The information in Table 1 will be extracted manually by KD or automatically by Endnote.
-1.
-The articles will be randomly selected for data extraction and two people (from MG, KD, BW and RR) will independently complete the data extraction form for each article.
+1. The articles will be randomly selected for data extraction and two people (from MG, KD, BW and RR) will independently complete the Data Extraction Form for each article. 
 Any conflicts will be discussed by the two extractors and resolved by mutual consent if possible. 
@@
 
-We will not attempt to retrieve any missing items in the data extraction form, in order to give an accurate assessment of the version of record and to avoid exhausting the extractors resources.
+We will not attempt to retrieve any missing items in the Data Extraction Form, in order to give an accurate assessment of the version of record and to avoid exhausting the extractors resources.

### 2020-01-01 qualtrics_forms/instructions.Rmd
rename forms file

--- 2019-11-15 data_extraction_form/instructions.Rmd
@@
 ---
-title: "Instructions: data extraction from epidemiology articles"
+title: "Preregistration: data extraction from epidemiology articles"
 author: "Katie Drax"
-date: "15 November 2019"
+date: "4 December 2019"
 output: word_document
@@
 
-- The study designs in **[number?]** articles have been identified by KD and MG. 
+- KD and MG independently identifed the study designs (cohort, case-control or cross-sectional) used by each of the 178 "observational epidemiology" articles by completing the Study Design Form for each article.
+In the form, extractors also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
+Any conflicts were discussed by the two extractors and resolved by mutual consent if possible. 
+If not resolved after discussion, RR determined the study designs. 
+The Study Design Form was hosted on Qualtrics but a PDF version can be found here https://osf.io/nd78g/.
 - Data extraction has not yet begun for any eligible articles
@@
 **[X] changes were made so the Data Extraction Form **[will/will not]** be completed again for the piloted articles.
+- KD assessed examined the notices attached to the online version of the observational epidemiology articles to identify any retracted articles. None were identified
 
-# Procedure
+# Data extraction procedure
 
@@
 
-1.
-Two people (from KD, RR, MG and BW) will independently identify the study design(s) (cohort, case-control or cross-sectional) used by each of the 178 articles by completing the Study Design Form for each article.
-The form also asks extractors to indicate if all supplementary material can be accessed via University of Bristol subscriptions. 
-We will attempt to retrieve any supplementary material we cannot access (e.g., by contacting the authors). 
-Any conflicts will be discussed by the two extractors and resolved by mutual consent if possible. 
-If not resolved after discussion, RR or MM will determine the study designs. 
-The Study Design Form was hosted on Qualtrics but a PDF version can be found here https://osf.io/nd78g/.  
+1. We will attempt to retrieve any supplementary material we could not access (e.g., by contacting the authors). 
 1.
 Articles containing multiple different study designs will be discussed on a case-by-case basis to confirm whether a single reporting checklist is adequate, or if multiple reporting checklists apply to a single article.
-1. KD will identify and exclude retracted articles by assessing the notices attached to the online version of the article.
 1. The information in Table 1 will be extracted manually by KD or automatically by Endnote.

### 2020-01-01 rscripts/clean_designs.R
make some checks reproducible and update comments

@@ -12,20 +12,21 @@ library(stringr)
 input <- "data/Classifying+epi+study+designs_11+December+2019_23.50.csv"
 # Import first three rows of the  Qualtrics csv export
 
-header <- read.csv(input, encoding = "UTF-8", nrows = 3)
+header <- read.csv(input, encoding = "UTF-8", nrows = 2, stringsAsFactors = F)
 
 # Row 1 should contain the "question numbers", row 2 the full question text and row 3 the "import id"s
 # As "Question Export Tags" assigned and question numbers changed to variable names the first row contains the correct variable names >
 # warn user to check row content and that variables in row 1 match full questions
 
-warning("manually check rows 1 = variables & row 2& 3 are import id and full questions")
-warning("manually check variables names match full questions")
+if(sum(str_count(colnames(header))) != 227) stop("check header colnames contains variable names")
+if(sum(grepl("ImportId", header[2,])) != 21) stop("check header row 2 contains import ids")
+if(sum(str_count(header[1, ])) != 1794) stop("check header row 1 contains full questions")
 
-# remove rows 2 and 3 by assigning column names (row 1)
+# assign column names to header
 
 header <- colnames(header)
 
-# remove row 2 and 3 on import skipping first three rows and assign header (row 1) as col.names
+# remove row 2 and 3 on import skipping first three rows and assign header as col.names
 
 df <- read.csv(input, skip = 3,
                encoding = "UTF-8", header = F, col.names = header, na.strings = c("", " "), stringsAsFactors = F)
@@ -38,6 +39,7 @@ df <- read.csv(input, skip = 3,
 
 qual_cols <- colnames(df)[1:10]
 qual_cols
+warning("check qual_cols above")
 
 # remove qual_cols
 
@@ -71,6 +73,7 @@ df$design_all <- gsub("no_statement,|NA,|,NA", "", df$design_all)
 df$design_all[grep("cross-sectional,cohort", df$design_all)] <- "cohort,cross-sectional"
 df$design_all[grep("cross-sectional,cross-sectional", df$design_all)] <- "cross-sectional"
 
+table(df$design)
 ####################
 # merge preparation ####
 ##################
@@ -86,21 +89,26 @@ articles <- read.csv("https://osf.io/8uy9w/?action=download", encoding = "UTF-8"
[…]

### 2020-01-03 rscripts/clean_designs.R
update input file to most recent, add checks, add title_sub column, set designs of articles without supp material as 'TBC' so will show up in conflicts, check & resolve duplicate titles & article ids, check all articles classified, output cleaned designs not conflicts

@@ -9,7 +9,7 @@ library(stringr)
 
 # assign input file 
 
-input <- "data/Classifying+epi+study+designs_11+December+2019_23.50.csv"
+input <- "data/Classifying+epi+study+designs_1+January+2020_21.04.csv"
 # Import first three rows of the  Qualtrics csv export
 
 header <- read.csv(input, encoding = "UTF-8", nrows = 2, stringsAsFactors = F)
@@ -31,15 +31,19 @@ header <- colnames(header)
 df <- read.csv(input, skip = 3,
                encoding = "UTF-8", header = F, col.names = header, na.strings = c("", " "), stringsAsFactors = F)
 
+# check length correct
+
+if(nrow(df) != 358) stop("too few/many observations")
+
 ############
 # clean ####
 ###########
 
-# vector of cols automatically outputted by Qualtrics (always first 10 cols if responses anonymised)
+# vector of cols automatically outputted by Qualtrics (always first 10 cols if responses anonymised and should be 112 characters)
 
 qual_cols <- colnames(df)[1:10]
-qual_cols
-warning("check qual_cols above")
+
+if(sum(str_count(qual_cols)) != 112) stop("check qual_cols contains qualtrics columns")
 
 # remove qual_cols
 
@@ -54,11 +58,35 @@ df <- df[order(df$article_id), ]
 
 df$initials <- tolower(df$initials)
 df$initials <- gsub("[[:punct:]]", "", df$initials)
-table(df$initials)
+if(sum(df$initials != "kd" |df$initials != "mg") != length(df$article_id)) stop("initials contains strings other than kd and mg")
+
+#################
+# check clean title ####
+##################
+
+# function to clean string 
+clean_string <- function(string){
+  # string: a vector of strings
+  # remove any non-english character, numbers, spaces or punctuation and lower
+  string <- gsub("[^\u0001-\u007F]+","", string)
+  string <- gsub("[0-9]", "", string)
+  string <- gsub("[[:punct:]]", "", string)
+  string <- tolower(string)
+  string <- gsub("[[:space:]]", "", string)
+}
+
+# column containing clean substring of title
+
+df$title_sub <- clean_string(df$title)
 
 #################
 # new column of designs ####
 ######################
+# some designs are missing because the supplmentary material is inaccessible material
+# set missing designs to to be confirmed (TBC)
+
[…]

### 2020-01-03 rscripts/resolve_conflicts.R
initial commit

@@ -1,0 +1,228 @@
+library(dplyr)
+
+both <- read.csv("outputs/clean_designs.csv", stringsAsFactors = F, encoding = "UTF-8")
+
+# function to check conflicts
+
+find_cons <- function(df,kd_col,mg_col){
+  # create df of katie's and Mark's columns that don't match
+  x <- df[which(kd_col != mg_col),]
+  # if conflicts return df and present warning
+  if(nrow(x) >0){
+    warning("conflicts")
+    return(x)
+  } else {
+    # if no conflicts print confirmation
+    print("no conflicts")
+  }
+}
+
+#############
+# get articles ####
+################
+
+# df of all articles in csv_clean_epi.csv on OSF
+
+articles_df <- read.csv("https://osf.io/8uy9w/?action=download", encoding = "UTF-8", stringsAsFactors = F)
+
+# vector of all article id's in csv_clean_epi.csv on oSF
+
+articles <- articles_df$id
+
+########################
+# find article_access conflicts ####
+#########################
+
+# find article access conflicts
+article_cons <- find_cons(both, both$access_article.kd, both$access_article.mg)%>%
+  select(c(article_id, access_article.kd, access_article.mg, title)) %>%
+  print()
+
+# mark could not find article for Tyree2015bank0732 so he incorrectly answered "Not present"
+# if article_cons only contain Tyree2015bank0732 correct Mark's responses to "Present but not accessible"
+if(article_cons$article_id == "Tyrre2015bank0732"){
+  both$access_article.mg[both$article_id == "Tyrre2015bank0732"] <- "Present but not accessible"
+} else {
+  stop("Tyree2015bank0732 not the only article access conflict")
+}
+
+# check article conflicts again
+
+check_art <- find_cons(both, both$access_article.kd, both$access_article.mg)
+
+if(check_art == "no conflicts"){
+  # if no conflicts remove one access_supp column
+  both$access_article.kd <- NULL
+  # rename the remaining column as access_supp
+  colnames(both)[colnames(both) == "access_article.mg"] <- "access_article"
+} else {
+  stop("still has access_supp conflicts")
+}
+
+###############################
+# find supp material (SM) access conflicts ####
+###############################
+
+# add column indicating if SM has been tracked down for articles with "Present but not availiable" SM
+
[…]

### 2020-01-03 rscripts/resolve_conflicts.R
add any non-missing values in Katie access_supp and access_article to Mark's columns so they won't be lost when Katie's columns are deleted

@@ -51,6 +51,8 @@ if(article_cons$article_id == "Tyrre2015bank0732"){
 check_art <- find_cons(both, both$access_article.kd, both$access_article.mg)
 
 if(check_art == "no conflicts"){
+  # add any access_article values in Katie's column that are missing in Marks so these aren't lost
+  both$access_article.mg[is.na(both$access_article.mg)] <- both$access_article.kd[is.na(both$access_article.mg)]
   # if no conflicts remove one access_supp column
   both$access_article.kd <- NULL
   # rename the remaining column as access_supp
@@ -107,7 +109,10 @@ both$access_supp.kd[both$article_id == "Cummi2016bank53-X"] <- "Present but not 
 check_supp <- find_cons(both, both$access_supp.kd, both$access_supp.mg)
 
 if(check_supp == "no conflicts"){
-  # if no conflicts remove one access_supp column
+  # if not conflicts remove katie's access_supp column
+  # add any access_supp values in Katie's column that are missing in Marks so these aren't lost
+  both$access_supp.mg[is.na(both$access_supp.mg)] <- both$access_supp.kd[is.na(both$access_supp.mg)]
+  # then remove Katies access_supp column
   both$access_supp.kd <- NULL
   # rename the remaining column as access_supp
   colnames(both)[colnames(both) == "access_supp.mg"] <- "access_supp"

### 2020-01-04 rscripts/resolve_conflicts.R
change 'both' to 'df', correct some checks, change NA to 'NA' so NA will appear as conflict if compared to another value, add design column

@@ -1,7 +1,8 @@
 library(dplyr)
 
-both <- read.csv("outputs/clean_designs.csv", stringsAsFactors = F, encoding = "UTF-8")
-
+df <- read.csv("outputs/clean_designs.csv", stringsAsFactors = F, encoding = "UTF-8")
+# set NA to actual value iso that it won't be ignored when matching
+df[is.na(df)] <- "NA"
 # function to check conflicts
 
 find_cons <- function(df,kd_col,mg_col){
@@ -23,7 +24,7 @@ find_cons <- function(df,kd_col,mg_col){
 
 # df of all articles in csv_clean_epi.csv on OSF
 
-articles_df <- read.csv("https://osf.io/8uy9w/?action=download", encoding = "UTF-8", stringsAsFactors = F)
+#articles_df <- read.csv("https://osf.io/8uy9w/?action=download", encoding = "UTF-8", stringsAsFactors = F)
 
 # vector of all article id's in csv_clean_epi.csv on oSF
 
@@ -34,31 +35,31 @@ articles <- articles_df$id
 #########################
 
 # find article access conflicts
-article_cons <- find_cons(both, both$access_article.kd, both$access_article.mg)%>%
+article_cons <- find_cons(df, df$access_article.kd, df$access_article.mg)%>%
   select(c(article_id, access_article.kd, access_article.mg, title)) %>%
   print()
 
 # mark could not find article for Tyree2015bank0732 so he incorrectly answered "Not present"
 # if article_cons only contain Tyree2015bank0732 correct Mark's responses to "Present but not accessible"
-if(article_cons$article_id == "Tyrre2015bank0732"){
-  both$access_article.mg[both$article_id == "Tyrre2015bank0732"] <- "Present but not accessible"
+if("Tyrre2015bank0732" %in% article_cons$article_id){
+  df$access_article.mg[df$article_id == "Tyrre2015bank0732"] <- "Present but not accessible"
 } else {
-  stop("Tyree2015bank0732 not the only article access conflict")
+  stop("Tyree2015bank0732 has no article access conflict")
 }
 
 # check article conflicts again
 
-check_art <- find_cons(both, both$access_article.kd, both$access_article.mg)
+check_art <- find_cons(df, df$access_article.kd, df$access_article.mg)
 
 if(check_art == "no conflicts"){
   # add any access_article values in Katie's column that are missing in Marks so these aren't lost
[…]

### 2020-01-05 rscripts/clean_designs.R
do not replace missing designs with TBC

@@ -82,11 +82,6 @@ df$title_sub <- clean_string(df$title)
 #################
 # new column of designs ####
 ######################
-# some designs are missing because the supplmentary material is inaccessible material
-# set missing designs to to be confirmed (TBC)
-
-df$design[df$access_article != "Yes" | df$access_supp == "Present but not accessible"] <- "TBC"
-  
 
 # new column combining design_judg and design to see stated and judged designs together
 
@@ -271,6 +266,18 @@ both$title_sub.kd <- NULL
 # rename remaining columns
 colnames(both)[colnames(both) == "title.mg"] <- "title"
 colnames(both)[colnames(both) == "title_sub.mg"] <- "title_sub"
+
+##################################
+# assign random number list ####
+##############################
+
+# if article ids are identical to those in csv_clean_epi add random number list
+
+if(identical(sort(both$article_id), sort(articles))){
+  set.seed(1)
+  both$num <- sample(1:length(both$article_id), length(both$article_id), replace = F)
+  both <- select(both, num, everything())
+}
 
 ############
 # export ####

### 2020-01-05 rscripts/resolve_conflicts.R
add TEMPORARY code to output csv of designs, create function to rename columns, change find_cons function so will find NA-not equal NA mismatches and rename dataframes containing conflicts

@@ -1,14 +1,16 @@
 library(dplyr)
 
 df <- read.csv("outputs/clean_designs.csv", stringsAsFactors = F, encoding = "UTF-8")
-# set NA to actual value iso that it won't be ignored when matching
-df[is.na(df)] <- "NA"
+
 # function to check conflicts
 
-find_cons <- function(df,kd_col,mg_col){
-  # create df of katie's and Mark's columns that don't match
-  x <- df[which(kd_col != mg_col),]
-  # if conflicts return df and present warning
+
+find_cons <- function(df,a,b){
+  # a & b = two columns to be compared
+  # df = dataframe columns are in
+  # create df of katie's and Mark's columns that don't match and include NAs in matching
+  x <- df[which((!is.na(a) & !is.na(b) & a==b | is.na(a) & is.na(b)) == F),]
+  # if conflicts - save x and present warning
   if(nrow(x) >0){
     warning("conflicts")
     return(x)
@@ -18,13 +20,30 @@ find_cons <- function(df,kd_col,mg_col){
   }
 }
 
+# merge and rename column if two columns identical
+rename_col <- function(df, a, b, b_new_name){
+  # a = column you want to delete
+  # b = column you want to keep and rename
+  # check if there's conflicts
+  if(identical(df[[a]], df[[b]])){
+    # remove a column
+    df[[a]] <- NULL
+    # rename b
+    colnames(df)[colnames(df) == b] <- b_new_name
+    return(df)
+    } else {
+      # if still conflicts stop merge
+      stop("still conflicts so can't merge columns")
+    }
+}
+
 #############
 # get articles ####
 ################
 
 # df of all articles in csv_clean_epi.csv on OSF
 
-#articles_df <- read.csv("https://osf.io/8uy9w/?action=download", encoding = "UTF-8", stringsAsFactors = F)
+articles_df <- read.csv("https://osf.io/8uy9w/?action=download", encoding = "UTF-8", stringsAsFactors = F)
 
 # vector of all article id's in csv_clean_epi.csv on oSF
 
@@ -35,32 +54,21 @@ articles <- articles_df$id
 #########################
 
 # find article access conflicts
-article_cons <- find_cons(df, df$access_article.kd, df$access_article.mg)%>%
+cons_art <- find_cons(df, df$access_article.kd, df$access_article.mg)%>%
   select(c(article_id, access_article.kd, access_article.mg, title)) %>%
   print()
 
[…]

### 2020-01-05 rscripts/clean_designs.R
correct checks, update input data, delete creation of design_all column (this was pasted into resolve_conflicts.R)

@@ -2,6 +2,7 @@
 
 library(dplyr)
 library(stringr)
+library(tidyr)
 
 ##########
 # import####
@@ -9,7 +10,7 @@ library(stringr)
 
 # assign input file 
 
-input <- "data/Classifying+epi+study+designs_1+January+2020_21.04.csv"
+input <- "data/Classifying+epi+study+designs_5+January+2020_19.38.csv"
 # Import first three rows of the  Qualtrics csv export
 
 header <- read.csv(input, encoding = "UTF-8", nrows = 2, stringsAsFactors = F)
@@ -33,7 +34,7 @@ df <- read.csv(input, skip = 3,
 
 # check length correct
 
-if(nrow(df) != 358) stop("too few/many observations")
+if(nrow(df) != 359) stop("too few/many observations")
 
 ############
 # clean ####
@@ -79,29 +80,41 @@ clean_string <- function(string){
 
 df$title_sub <- clean_string(df$title)
 
-#################
-# new column of designs ####
-######################
+################
+# check designs ####
+##############
 
-# new column combining design_judg and design to see stated and judged designs together
+# check designs only missing because material was inaccessible
 
-df$design_all <- paste(df$design, df$design_judg, sep = ",")
+if(sum(df$access_article[is.na(df$design)] != "Yes" | df$access_supp[is.na(df$design)] == "Present but not accessible") != sum(is.na(df$design))){
+  stop("check NA designs")
+} else {
+  print("all NA designs had inaccesible material")
+}
 
-# remove NAs added by paste 
+# check all design responses are correct
 
-df$design_all[df$design_all == "NA,NA"] <- NA
-df$design_all <- gsub("no_statement,|NA,|,NA", "", df$design_all)
+design_labels <- c("case-control", "no_statement", "cross-sectional", "cohort", "cohort,cross-sectional", "cohort,case-control", "cross-sectional,no_statement")
 
-# standarise order of designs
-df$design_all[grep("cross-sectional,cohort", df$design_all)] <- "cohort,cross-sectional"
-df$design_all[grep("cross-sectional,cross-sectional", df$design_all)] <- "cross-sectional"
+if(length(setdiff(df$design[!is.na(df$design)], design_labels)) == 0){
+  print("correct design responses")
+} else{
+  print(setdiff(df$design[!is.na(df$design)], design_labels))
+  stop("incorrect design responses")
+}
 
-# check designs
+# check all design_judg responses are correct
 
[…]

### 2020-01-05 rscripts/resolve_conflicts.R
added all temporary code (e.g. setting conflicted designs to TBC and exporting list of conficts for becky) to temporary.R, add code exporting unresolved_conflicts if conflicts remain and resolved_conflicts if no conflicts, paste in code creating design_all column from clean_designs.R

@@ -1,11 +1,21 @@
 library(dplyr)
 
+##########
+# import ####
+##########
+
 df <- read.csv("outputs/clean_designs.csv", stringsAsFactors = F, encoding = "UTF-8")
+
+##############
+# functions ####
+#############
 
 # function to check conflicts
 
 
 find_cons <- function(df,a,b){
+  if(is.null(a)) stop("a doesn't exist")
+  if(is.null(b)) stop("b doesn't exist")
   # a & b = two columns to be compared
   # df = dataframe columns are in
   # create df of katie's and Mark's columns that don't match and include NAs in matching
@@ -36,6 +46,51 @@ rename_col <- function(df, a, b, b_new_name){
       stop("still conflicts so can't merge columns")
     }
 }
+
+
+#################
+# new column of designs ####
+######################
+
+# new column combining design_judg and design to see stated and judged designs together
+
+df$design_all.kd <- paste(df$design.kd, df$design_judg.kd, sep = ",")
+df$design_all.mg <- paste(df$design.mg, df$design_judg.mg, sep = ",")
+# remove NAs added by paste 
+
+df$design_all.kd[df$design_all.kd == "NA,NA"] <- NA
+df$design_all.mg[df$design_all.mg == "NA,NA"]  <- NA
+
+df$design_all.kd <- gsub("no_statement,|NA,|,NA", "", df$design_all.kd)
+df$design_all.mg <- gsub("no_statement,|NA,|,NA", "", df$design_all.mg)
+
+# standarise order of designs
+df$design_all.kd[grep("cross-sectional,cohort", df$design_all.kd)] <- "cohort,cross-sectional"
+df$design_all.kd[grep("cross-sectional,cross-sectional", df$design_all.kd)] <- "cross-sectional"
+df$design_all.mg[grep("cross-sectional,cohort", df$design_all.mg)] <- "cohort,cross-sectional"
+
+# check all design responses are correct
+
+design_all_labels <- c("case-control", "other", "no_statement", "cross-sectional", "cohort", 
+                       "cohort,cross-sectional", "cohort,case-control", "cross-sectional,no_statement", "cross-sectional,other")
+
+if(length(setdiff(df$design_all.kd[!is.na(df$design_all.kd)], design_all_labels)) == 0){
+  print("correct design_all.kd responses")
+} else{
+  print(setdiff(df$design_all.kd[!is.na(df$design_all.kd)], design_all_labels))
[…]

### 2020-01-05 rscripts/temporary.R
temp code creating temp files

@@ -1,0 +1,114 @@
+library(dplyr)
+
+##########
+# import ####
+##########
+
+temp <- read.csv("outputs/unresolved_designs.csv", stringsAsFactors = F, encoding = "UTF-8")
+
+##############
+# functions ####
+#############
+
+# function to check conflicts
+
+
+find_cons <- function(df,a,b){
+  if(is.null(a)) stop("a doesn't exist")
+  if(is.null(b)) stop("b doesn't exist")
+  # a & b = two columns to be compared
+  # df = dataframe columns are in
+  # create df of katie's and Mark's columns that don't match and include NAs in matching
+  x <- df[which((!is.na(a) & !is.na(b) & a==b | is.na(a) & is.na(b)) == F),]
+  # if conflicts - save x and present warning
+  if(nrow(x) >0){
+    warning("conflicts")
+    return(x)
+  } else {
+    # if no conflicts print confirmation
+    print("no conflicts")
+  }
+}
+
+# merge and rename column if two columns identical
+rename_col <- function(df, a, b, b_new_name){
+  # a = column you want to delete
+  # b = column you want to keep and rename
+  # check if there's conflicts
+  if(identical(df[[a]], df[[b]])){
+    # remove a column
+    df[[a]] <- NULL
+    # rename b
+    colnames(df)[colnames(df) == b] <- b_new_name
+    return(df)
+  } else {
+    # if still conflicts stop merge
+    stop("still conflicts so can't merge columns")
+  }
+}
+
+
+
+#####################
+# sort conflicts ####
+########################
+
+# find conflicts again
+cons_design <- find_cons(temp, temp$design_all.kd, temp$design_all.mg)
+cons_supp <- find_cons(temp, temp$access_supp.kd, temp$access_supp.mg)
+cons_art <- find_cons(temp, temp$access_article.kd, temp$access_article.mg)
+
+# set all designs of articles with some inaccessible materials to TBC
+
+temp$design_all.kd[(temp$access_article.kd != "Yes" & temp$access_article.mg != "Yes") | 
+                     (temp$access_supp.kd == "Present but not accessible" & temp$access_supp.mg == "Present but not accessible")] <- "TBC"
+
+temp$design_all.mg[(temp$access_article.kd != "Yes" & temp$access_article.mg != "Yes") | 
[…]

### 2020-01-06 rscripts/clean_designs.R
create design_all column in clean

@@ -116,6 +116,37 @@ if(length(setdiff(df$design_judg[!is.na(df$design_judg)], design_judg_labels)) =
   stop("incorrect design_judg responses")
 }
 
+#################
+# column of all designs ####
+######################
+
+# new column combining design_judg and design to see stated and judged designs together
+
+df$design_all <- paste(df$design, df$design_judg, sep = ",")
+
+# remove NAs added by paste 
+
+df$design_all[df$design_all == "NA,NA"] <- NA
+
+df$design_all <- gsub("no_statement,|NA,|,NA", "", df$design_all)
+
+# standarise order of designs
+df$design_all[grep("cross-sectional,cohort", df$design_all)] <- "cohort,cross-sectional"
+df$design_all[grep("cross-sectional,cross-sectional", df$design_all)] <- "cross-sectional"
+
+# check all design responses are correct
+
+design_all_labels <- c("case-control", "other", "no_statement", "cross-sectional", "cohort", 
+                       "cohort,cross-sectional", "cohort,case-control", "cross-sectional,no_statement", "cross-sectional,other")
+
+if(length(setdiff(df$design_all[!is.na(df$design_all)], design_all_labels)) == 0){
+  print("correct design_all responses")
+} else{
+  print(setdiff(df$design_all[!is.na(df$design_all)], design_all_labels))
+  stop("incorrect design_all responses")
+}
+
+
 ####################
 # merge preparation ####
 ##################

### 2020-01-06 rscripts/resolve_conflicts.R

@@ -47,51 +47,6 @@ rename_col <- function(df, a, b, b_new_name){
     }
 }
 
-
-#################
-# new column of designs ####
-######################
-
-# new column combining design_judg and design to see stated and judged designs together
-
-df$design_all.kd <- paste(df$design.kd, df$design_judg.kd, sep = ",")
-df$design_all.mg <- paste(df$design.mg, df$design_judg.mg, sep = ",")
-# remove NAs added by paste 
-
-df$design_all.kd[df$design_all.kd == "NA,NA"] <- NA
-df$design_all.mg[df$design_all.mg == "NA,NA"]  <- NA
-
-df$design_all.kd <- gsub("no_statement,|NA,|,NA", "", df$design_all.kd)
-df$design_all.mg <- gsub("no_statement,|NA,|,NA", "", df$design_all.mg)
-
-# standarise order of designs
-df$design_all.kd[grep("cross-sectional,cohort", df$design_all.kd)] <- "cohort,cross-sectional"
-df$design_all.kd[grep("cross-sectional,cross-sectional", df$design_all.kd)] <- "cross-sectional"
-df$design_all.mg[grep("cross-sectional,cohort", df$design_all.mg)] <- "cohort,cross-sectional"
-
-# check all design responses are correct
-
-design_all_labels <- c("case-control", "other", "no_statement", "cross-sectional", "cohort", 
-                       "cohort,cross-sectional", "cohort,case-control", "cross-sectional,no_statement", "cross-sectional,other")
-
-if(length(setdiff(df$design_all.kd[!is.na(df$design_all.kd)], design_all_labels)) == 0){
-  print("correct design_all.kd responses")
-} else{
-  print(setdiff(df$design_all.kd[!is.na(df$design_all.kd)], design_all_labels))
-  stop("incorrect design_all.kd responses")
-}
-
-if(length(setdiff(df$design_all.mg[!is.na(df$design_all.mg)], design_all_labels)) == 0){
-  print("correct design_all.mg responses")
-} else{
-  print(setdiff(df$design_all.mg[!is.na(df$design_all.mg)], design_all_labels))
-  stop("incorrect design_all.mg responses")
-}
-
-# drop unnecessary columns
-
-df <- select(df, -c(design.kd, design.mg, design_judg.kd, design_judg.mg))
-
 #############
 # get articles ####
 ################

### 2020-01-06 rscripts/clean_designs.R
move function up and sort by random number

@@ -1,8 +1,24 @@
+#############
 # libraries ####
+#############
 
 library(dplyr)
 library(stringr)
-library(tidyr)
+
+###############
+# functions ####
+##############
+
+# function to clean string 
+clean_string <- function(string){
+  # string: a vector of strings
+  # remove any non-english character, numbers, spaces or punctuation and lower
+  string <- gsub("[^\u0001-\u007F]+","", string)
+  string <- gsub("[0-9]", "", string)
+  string <- gsub("[[:punct:]]", "", string)
+  string <- tolower(string)
+  string <- gsub("[[:space:]]", "", string)
+}
 
 ##########
 # import####
@@ -54,29 +70,13 @@ df <- select(df, -qual_cols)
 
 df <- df[order(df$article_id), ]
 
-
 # standardise initials by lowering and removing punctuation
 
 df$initials <- tolower(df$initials)
 df$initials <- gsub("[[:punct:]]", "", df$initials)
 if(sum(df$initials != "kd" |df$initials != "mg") != length(df$article_id)) stop("initials contains strings other than kd and mg")
 
-#################
-# check clean title ####
-##################
-
-# function to clean string 
-clean_string <- function(string){
-  # string: a vector of strings
-  # remove any non-english character, numbers, spaces or punctuation and lower
-  string <- gsub("[^\u0001-\u007F]+","", string)
-  string <- gsub("[0-9]", "", string)
-  string <- gsub("[[:punct:]]", "", string)
-  string <- tolower(string)
-  string <- gsub("[[:space:]]", "", string)
-}
-
-# column containing clean substring of title
+# column containing clean substring of title for easier matching
 
 df$title_sub <- clean_string(df$title)
 
@@ -319,5 +319,9 @@ if(identical(sort(both$article_id), sort(articles))){
 ############
 # export ####
 #############
+
+#sort by random number
+
+both <- both[order(both$num), ]
 
 write.csv(both, "outputs/clean_designs.csv", row.names = F, fileEncoding = "UTF-8")

### 2020-01-06 rscripts/resolve_conflicts.R
better export check and set articles with inaccessible material to TBC designs

@@ -223,6 +223,14 @@ df$resol_becky[df$article_id == "Sarka2018ants.009"] <- T
 df$design_all.kd[!is.na(df$resol_correct_design)] <- df$resol_correct_design[!is.na(df$resol_correct_design)]
 df$design_all.mg[!is.na(df$resol_correct_design)] <- df$resol_correct_design[!is.na(df$resol_correct_design)]
 
+# set all designs of articles with some inaccessible materials to TBC
+
+df$design_all.kd[df$access_article != "Yes" | df$access_supp == "Present but not accessible"] <- "TBC"
+
+df$design_all.mg[df$access_article != "Yes" | df$access_supp == "Present but not accessible"] <- "TBC"
+
+
+
 # if no conflicts rename design column
 
 df <- rename_col(df, "design_all.kd", "design_all.mg", "design")
@@ -242,9 +250,9 @@ cons_art <- find_cons(df, df$access_article.kd, df$access_article.mg)
 
 #  export
 
-if(nrow(cons_design) != 0){
+if(exists("cons_design")|exists("cons_supp") | exists("cons_art")|sum(is.na(df$design)) >0|length(df$design[df$design == "TBC"]) > 0){
   # if conflicts export df as is
-  warning("still conflicts so exporting unresolved designs")
+  warning("still conflicts, NA designs or TBC designs so exported unresolved designs")
   # export df as is
   write.csv(df, "outputs/unresolved_designs.csv", row.names = F, fileEncoding = "UTF-8")
   } else {

### 2020-01-06 rscripts/check_extraction_form.R
update

@@ -7,50 +7,46 @@
 library(dplyr)
 
 # skip question name row so question row will be header
-df <- read.csv("data/Data+Extraction+Form_2+December+2019_17.58.csv", skip = 1)
+df <- read.csv("qualtrics_forms/Data+Extraction+Form_6+January+2020_17.10.csv", 
+               skip = 1, encoding = "UTF-8", stringsAsFactors = F)
 
-# save column names from raw df
 
-df_names <- colnames(df)
-
-#### remove unnecessary cols ####
-
-# vector of columns qualtrics automatically exports
-qual_cols <- colnames(df)[1:17]
+# vector of columns qualtrics automatically exports (always first 10 if response autonomised)
+qual_cols <- colnames(df)[1:10]
 
 # remove qual_cols
 
 df <- select(df, -qual_cols)
 
-#### remove rows ####
+# find design specific cols
+
+design_qs <- grepl("sectional|case.control|cohort", colnames(df), ignore.case = T)
+
+# variable name is in question text so remove question text from colnames that is not the name
+colnames(df) <- tolower(colnames(df))
+num <- grepl("\\d", colnames(df))
+colnames(df)[num] <- gsub("^[^\\d]+", "", colnames(df)[num], perl = T)
+colnames(df) <- gsub("(?<=\\.)(.*?)(?=text)", "", colnames(df), perl = T)
+colnames(df) <- gsub("\\.\\..*$", "", colnames(df), perl = T)
+
 
 # remove row automatically outputted by qualtrics
-id <- grep("import", df$Article.ID...copy.and.paste.in.the.article.ID.from.the.csv.file..close.Excel, ignore.case = T)
-df <- df[-1, ]
-
-# remove real responses 
-
-#### separate out strobe it
-id <- grep("^[a-zA-Z]{0,6}$", df$Your.initials)
+id <- grep("import", df$article_id, ignore.case = T)
 df <- df[-id, ]
 
-#### remove cols not needed for test ####
+
+#### remove cols not needed for design specific test ####
 
 # remove "Text" cols
-text <- grep("Text", colnames(df))
+text <- grep(".Text", colnames(df))
 
 test_df <- df[, -text]
 
 # remove cols that aren't design specific
-design_qs <- grep("sectional|case.control|cohort", colnames(test_df), ignore.case = T)
 
 test_df <- test_df[, design_qs]
 
[…]

### 2020-01-06 rscripts/temporary.R

@@ -58,44 +58,63 @@ cons_design <- find_cons(temp, temp$design_all.kd, temp$design_all.mg)
 cons_supp <- find_cons(temp, temp$access_supp.kd, temp$access_supp.mg)
 cons_art <- find_cons(temp, temp$access_article.kd, temp$access_article.mg)
 
-# set all designs of articles with some inaccessible materials to TBC
-
-temp$design_all.kd[(temp$access_article.kd != "Yes" & temp$access_article.mg != "Yes") | 
-                     (temp$access_supp.kd == "Present but not accessible" & temp$access_supp.mg == "Present but not accessible")] <- "TBC"
-
-temp$design_all.mg[(temp$access_article.kd != "Yes" & temp$access_article.mg != "Yes") | 
-                     (temp$access_supp.kd == "Present but not accessible" & temp$access_supp.mg == "Present but not accessible")] <- "TBC"
-
-# set all designs of articles currently in conflict to  TBC
-
-temp$design_all.kd[temp$article_id %in% cons_design$article_id] <- "TBC"
-temp$design_all.mg[temp$article_id %in% cons_design$article_id] <- "TBC"
-
-# try to rename again
-
-temp <- rename_col(temp, "access_article.kd", "access_article.mg", "access_article")
-temp <- rename_col(temp, "access_supp.kd", "access_supp.mg", "access_supp")
-temp <- rename_col(temp, "design_all.kd", "design_all.mg", "design")
-
-#  no conflicts remain so don't need any columns that were used to resolve conflicts
-
-cols <- colnames(temp)[grepl("comments|design_|resol|design.|title_sub", colnames(temp)) == F]
-temp <- temp[, colnames(temp)%in% cols]
-
-# order by number 
-temp <- temp[order(temp$num),]
 ###########################
-# export first 80 articles ####
+# get first 80 articles ####
 ############################
 
-first_lot <- temp[temp$num %in% 1:80, ]
+if(exists("cons_design")|exists("cons_supp") | exists("cons_art")|sum(is.na(temp$design)) != 0){
+  stop("still conflicts and missing designs")
+} else {
+  #no conflicts remain so don't need any columns that were used to resolve conflicts
+  cols <- colnames(temp)[grepl("comments|design_|resol|design.|title_sub", colnames(temp)) == F]
+  # select first 80 random numbers
[…]

### 2020-01-07 rscripts/check_extraction_form.R
update

@@ -8,12 +8,16 @@ library(dplyr)
 
 # skip question name row so question row will be header
 df <- read.csv("qualtrics_forms/Data+Extraction+Form_6+January+2020_17.10.csv", 
-               skip = 1, encoding = "UTF-8", stringsAsFactors = F)
+               skip = 1, encoding = "UTF-8", stringsAsFactors = F, na.strings = c("", " "))
 
 
 # vector of columns qualtrics automatically exports (always first 10 if response autonomised)
 qual_cols <- colnames(df)[1:10]
 
+# rename article id
+
+colnames(df)[grep("article_id", colnames(df))] <- "article_id"
+colnames(df)[grep("^designs", colnames(df))] <- "designs"
 # remove qual_cols
 
 df <- select(df, -qual_cols)
@@ -22,46 +26,61 @@ df <- select(df, -qual_cols)
 
 design_qs <- grepl("sectional|case.control|cohort", colnames(df), ignore.case = T)
 
+# remove row automatically outputted by qualtrics
+id <- grep("import", df$article_id, ignore.case = T)
+df <- df[-id, ]
+
+#### remove cols not needed for design specific test ####
+
+# remove "Text" cols
+text <- grep("text", colnames(df), ignore.case = T)
+
+# remove cols that aren't design specific
+
+test_df <- df[, design_qs]
+
+# cols for cohort & cross-sectional designs ####
+
+# check  cols only contain questions for cohort and cross-sectional and no other two designs
+length(grep("^(?=.*_coh)(?=.*_cs).*$", colnames(df), ignore.case = T, perl = T) >0)
+length(grep("^(?=.*_coh)(?=.*_cc).*$", colnames(df), ignore.case = T, perl = T) >0)
+length(grep("^(?=.*_cs)(?=.*_cc).*$", colnames(df), ignore.case = T, perl = T) >0)
+
+# select columns for cross-sectional and cohort designs
+
+coh_cs <- colnames(df)[grep("^(?=.*_coh)(?=.*_cs).*$", colnames(df), ignore.case = T, perl = T)]
+cs <- grep("_cs", colnames(df)[-coh_x], ignore.case = T, value =T)
+cc <- grep("_cc", colnames(df), ignore.case = T, value =T)
+coh <- grep("_coh", colnames(df)[-coh_x], ignore.case = T, value = T)
+
[…]

### 2020-01-07 qualtrics_forms/instructions.Rmd
minor update

@@
 If not resolved after discussion, RR determined the study designs. 
-The Study Design Form was hosted on Qualtrics but a PDF version can be found here https://osf.io/nd78g/.
+The Study Design Form was hosted on Qualtrics but a PDF version can be found here **[]**.
 - Data extraction has not yet begun for any eligible articles
-- The Data Extraction form has been piloted on **[number]** observational epidemiology articles by MG, KD and RR. 
-**[X] changes were made so the Data Extraction Form **[will/will not]** be completed again for the piloted articles.
+- The Data Extraction form has been piloted on 5 observational epidemiology articles by MG, KD and RR.
+Major changes were made so the Data Extraction Form will be completed again for the piloted articles.
 - KD assessed examined the notices attached to the online version of the observational epidemiology articles to identify any retracted articles. None were identified
@@
 | Issue |   | Endnote  |
-| Open Access |   | (1) Search for article on open access button ([https://openaccessbutton.org/](https://openaccessbutton.org/)) and follow the link on non-university network  (2) If no on open access button check if there are links on Google Scholar |
-| Countries of corresponding, first &amp; last author |   | Extract country of author's affiliations |
-| UK Biobank Application number | Study's approved UK Biobank data application number (if present) | (1) Search article for all these terms:  'application' , 'approv' , 'number' , 'Resource' (2) If not found examine any strings of numbers in article |
-| Author keywords |   | (1) Extract from page 1 (2) If not on page 1 search article for 'keywords' |
-| Journal endorsement | Journal's endorsement of STROBE before publication | (1) Extract references to STROBE from the journal's most recent 'instructions/guide for authors' before article's publication date |
+| Open Access |   | See if can access article on publishing journal's site on a non-university network. If not accessible search for self-archived version on Google Scholar |
+| Countries of corresponding author |   | Extract country from author's affiliations |
+| UK Biobank Application number | Study's approved UK Biobank data application number (if present) | Search article for all these terms:  'application' , 'approv' , 'number' , 'Resource'.
+If not found examine any strings of numbers in article |
+| Author keywords |   | Search page 1. If not on page 1 search article for 'keywords' |
+| Journal endorsement | Journal's endorsement of STROBE before publication | Extract references to STROBE from the journal's most recent 'instructions/guide for authors' before article's publication date |
 | Journal impact factor |   | Retrieve the impact factor of the journal for year of publication from Web of Science's Journal Citation Reports (JCR) |
-| Corresponding author email address |   | Endnote; if missing: (1) extract from page 1 (2) If not on page 1 search article for 'corresponding', '@') |
-| Conflicts of interest statement |   | (1) Search article material for all these terms: 'competing', 'interest' ,'conflict' , 'disclosure' (2) If not found examine article for similar terms (3) if present extract conflicts of interest statements |
+| Corresponding author email address |   | Endnote; if missing: search article for 'corresponding', '@') |
+| Conflicts of interest statement |   | Search article material for all these terms: 'competing', 'interest' ,'conflict' , 'disclosure'grcy |
 | Classification | Classification ('mendelian randomisation', 'observational epidemiology' or 'other') as assigned by KD, BW and RR in Rayyan | Rayyan |

### 2020-01-09 qualtrics_forms/prereg_data_extraction.Rmd
rename, add exclusion of 1 article, remove bullets as this won't appear in OSF formatting

---
title: "Preregistration: data extraction from epidemiology articles"
author: "Katie Drax"
date: "9 January 2020"
output: word_document
---

# Time of writing

KD and MG independently identified the study designs (cohort, case-control or cross-sectional) used by each of the 178 "observational epidemiology" articles by completing the Study Design Form for each article. 
In the form, extractors also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
Any conflicts were discussed by the two extractors and resolved by mutual consent if possible. 
If not resolved after discussion, RR determined the study designs. 
The Study Design Form was hosted on Qualtrics but a PDF version can be found at https://osf.io/nd78g/

The articles were randomly assigned an id number from 1-178.
The articles with an id number between 1 and 80 were selected for the first round of data extraction.
KD was assigned as the first extractor to all 80 articles.
The second extractor was randomly assigned from MG, RR or BW.

The Data Extraction form was hosted on Qualtrics and piloted twice.
In the first, 5 observational epidemiology articles  were piloted by MG, KD and/or RR. 
Major changes were made so the Data Extraction Form will be completed again for the articles in the first pilot.
In the second, KD piloted the changed form on 3 articles.
No changes were made so the responses for articles in the second pilot will be retained.

KD assessed examined the notices attached to the online version of the observational epidemiology articles to identify any retracted articles. None were identified

1 article was incorrectly classified as an 'observational epidemiology' study so 177 observational epidemiology articles remain.

# Data extraction procedure

KD will attempt to retrieve any supplementary material we could not access by contacting the authors.

The information in Table 1 (attached) will be extracted manually by KD or automatically by Endnote which contains the article's metadata exported from the database the article was located in.

The extractors will independently complete the Data Extraction Form for each article they are assigned. 
Any conflicts will be discussed by the two extractors and resolved by mutual consent if possible. 
If not resolved after discussion, RR or MM will resolve the conflict. 
Articles will continue to be selected at random until none remain, or until the authors resources are exhausted.

*Table 1. Methods for extracting article meta-data*

| Item | Definition | Method of extraction |
| --- | --- | --- |
| Article ID(s) of related articles | Article IDs of articles that contain similar analyses. | Examine articles with same first author, last author and UK Biobank application number. |
| Citation |   | Endnote |
| Publication date |   | Endnote |
| Authors |   |  Endnote |
| Journal |   | Endnote  |
| Volume |   | Endnote  |
| Issue |   | Endnote  |
| Open Access |   | See if can access article on publishing journal's site on a non-university network. If not accessible search for self-archived version on Google Scholar |
| Countries of corresponding author |   | Extract country from author's affiliations |
| UK Biobank Application number | Study's approved UK Biobank data application number (if present) | Search article for all these terms:  'application' , 'approv' , 'number' , 'Resource'. If not found examine any strings of numbers in article |
| Author keywords |   | Search page 1. If not on page 1 search article for 'keywords' |
| Journal endorsement | Journal's endorsement of STROBE before publication | Extract references to STROBE from the journal's most recent 'instructions/guide for authors' before article's publication date |
| Journal impact factor |   | Retrieve the impact factor of the journal for year of publication from Web of Science's Journal Citation Reports (JCR) |
| Corresponding author email address |   | Search article for 'corresponding', '@') |
| Conflicts of interest statement |   | Search article material for all these terms: 'competing', 'interest' ,'conflict' , 'disclosure' |
| Classification | Classification ('mendelian randomisation', 'observational epidemiology' or 'other') as assigned by KD, BW and RR in Rayyan | Rayyan

# Changes from the preregistered protocol

We will not attempt to retrieve any missing items in the Data Extraction Form, in order to give an accurate assessment of the version of record and to avoid exhausting the extractors resources.

### 2020-01-09 rscripts/temporary.R
exlude not epi articles and output csv of these exluded articles

@@ -58,6 +58,17 @@ cons_design <- find_cons(temp, temp$design_all.kd, temp$design_all.mg)
 cons_supp <- find_cons(temp, temp$access_supp.kd, temp$access_supp.mg)
 cons_art <- find_cons(temp, temp$access_article.kd, temp$access_article.mg)
 
+#########
+# exclude ####
+#######
+
+# exclude any designs that need excluding because they are not epidemiological articles
+
+nepi <- grep("exclude_nepi", temp$design)
+temp <- temp[-nepi, ]
+# export list of not epi articles
+write.csv(temp[nepi, ], "outputs/not_epi.csv", fileEncoding = "UTF-8", row.names = F)
+
 ###########################
 # get first 80 articles ####
 ############################

### 2020-01-22 rscripts/jif_merge.R
initial commit

@@ -1,0 +1,39 @@
+# added year column to all journals 
+library(dplyr)
+
+add_year <- function(file, pattern, year){
+  source <- read.csv(file, stringsAsFactors = F, header = F, nrows = 1, na.strings = c(" ", "")) %>%
+    .[1, 1]
+  x <- read.csv(file, stringsAsFactors = F, skip = 1, na.strings = c(" ", ""))
+  if(grepl(pattern, source) & is.data.frame(x)){
+    x$year <- rep(year, nrow(x))
+    } else {
+      stop("not dataframe")
+      }
+  x <- x[!duplicated(x),]
+  if(length(grep("Copyright|you agree to the", x$Rank) != 2)){
+    text <- grep("Copyright|you agree to the", x$Rank)
+    x <- x[-text, ]
+    } else {
+      stop("no Rank values contain copyright and/or terms of use text")
+      }
+    return(x)
+    }
+
+jif12 <- add_year("data/jif/jif_2012.csv", "Year: 2012", 2012)
+jif13 <- add_year("data/jif/jif_2013.csv", "Year: 2013", 2013)
+jif14 <- add_year("data/jif/jif_2014.csv", "Year: 2014", 2014)
+jif15 <- add_year("data/jif/jif_2015.csv", "Year: 2015", 2015)
+jif16 <- add_year("data/jif/jif_2016.csv", "Year: 2016", 2016)
+jif17 <- add_year("data/jif/jif_2017.csv", "Year: 2017", 2017)
+jif18 <- add_year("data/jif/jif_2018.csv", "Year: 2018", 2018)
+
+all <- rbind(jif12, jif13, jif14, jif15, jif16, jif17, jif18)
+all$Full.Journal.Title <- tolower(all$Full.Journal.Title)
+all$JCR.Abbreviated.Title <- tolower(all$JCR.Abbreviated.Title)
+
+first <- read.csv("outputs/first_80.csv") %>%
+  select(., c("journal", "year"))%>%
+  .[!duplicated(.),]
+
+journals <- left_join(journals, all, by = c("journal" = "JCR.Abbreviated.Title", "year"))

### 2020-01-22 qualtrics_forms/prereg_data_extraction.Rmd
final prereg version and rename qualtrics forms

@@
 date: "9 January 2020"
-output: word_document
+output: pdf_document
 ---
@@
 
-The information in Table 1 (attached) will be extracted manually by KD or automatically by Endnote which contains the article's metadata exported from the database the article was located in.
+The items in Table 1 will be extracted manually by KD or automatically by Endnote which contains the article's metadata exported from the database the article was located in.
 
@@
 | Open Access |   | See if can access article on publishing journal's site on a non-university network. If not accessible search for self-archived version on Google Scholar |
-| Countries of corresponding author |   | Extract country from author's affiliations |
+| Country | Countries of corresponding author | Extract country from corresponding author's affiliations |
 | UK Biobank Application number | Study's approved UK Biobank data application number (if present) | Search article for all these terms:  'application' , 'approv' , 'number' , 'Resource'.
@@
 | Journal endorsement | Journal's endorsement of STROBE before publication | Extract references to STROBE from the journal's most recent 'instructions/guide for authors' before article's publication date |
-| Journal impact factor |   | Retrieve the impact factor of the journal for year of publication from Web of Science's Journal Citation Reports (JCR) |
-| Corresponding author email address |   | Search article for 'corresponding', '@') |
+| Journal impact factor |   | Retrieve the impact factor of the journal for year of publication from Web of Science's Journal Citation Reports |
+| Corresponding author email address |   | Search article for 'corresponding', '@' |
 | Conflicts of interest statement |   | Search article material for all these terms: 'competing', 'interest' ,'conflict' , 'disclosure' |
-| Classification | Classification ('mendelian randomisation', 'observational epidemiology' or 'other') as assigned by KD, BW and RR in Rayyan | Rayyan
+| Classification | Classification assigned by KD, BW and RR in Rayyan | Rayyan
 

### 2021-06-03 manuscript/manuscript.Rmd
add outputs folder, manuscript folder, and field notes

---
title: "The UK Biobank Study: Systematic Review of the Reporting Quality in UK Biobank Studies"
author: "Emma Anderson, George Davey Smith, Katie Drax, Mark Gibson, Marcus Munafò, Benjamin Woolf, and Rebecca Richmond"
date: "`r format(Sys.time(), '%d %B %Y')`"
output: 
  word_document:
      number_sections: true
bibliography: manuscript.bib
csl: apa.csl
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Background

There is a general move towards making data freely accessible, or ‘open’, by academic and governmental institutions alike [@vasilevsky2017; @worldwidewebfoundation2018]. 
Completely open data allows any individual, regardless of their motivation, discipline or skillset, to access and analyse a dataset for free. 
This accessibility offers researchers new opportunities but there is also uncertainty about how the nature of open data may impact the reliability of research findings. 
One method by which the reliability of results from open data may be investigated is through replication. 
Direct replication of secondary data analyses is aided by the publication of an original analysis plan, in which the authors describe their statistical analysis in full. 
However, recent evidence suggests that it is unclear whether this is sufficient to ensure direct replicability of findings [@hardwicke2018; @naudet2018].

To investigate the reliability of findings from open data, we will attempt to replicate all eligible research articles that use UK Biobank (UKB) data. 
The UKB is a prospective cohort study which collected a wide range of health data from 500,000 people aged 40-69 in the UK during 2006-2010 [@ukbiobankcoordinatingcentre2007].
Importantly, UKB data is partially open data because it is only available to bona fide researchers conducting research in the public interest on a cost-recovery basis.
The UKB reserves the right to reject access applications for projects it deems inappropriate [@ukbiobankcoordinatingcentre2011].
We first conducted a systematic review to (1) assess the completeness of UKB study reporting, (2) identify articles that are suitable for replication, and (3) inform reporting guidelines for other articles using the UKB and similar resources. 
We will evaluate reproducibility of studies conducted in the UKB given the abundance of published articles using data from this cohort and their contemporary nature, increasing the likelihood that authors will be contactable.
Furthermore, they all use the same, professionally collected, maintained and diverse open resource. 
This guarantees data accessibility and usability and means that differences between the original and replication results cannot be explained by poor data curation or heterogeneity between studies.

Many previous systematic reviews of reporting standards across a range of disciplines used guidelines to assess the completeness of reporting [e.g. @agha2014; @cook2011; @plint2006).
We used the STROBE (STrengthening the Reporting of Observational Studies in Epidemiology) Statement, which provides reporting guidelines for observational studies, to evaluate articles arising from UKB and to make comparisons between the articles. 
The STROBE Statement has been endorsed by over 100 journals and, as no major changes have been made to the original version published in 2007 [@cevallos2014], these guidelines are relevant to all studies conducted using the UKB.
Given that studies conducted using UKB may be cohort, cross-sectional or ‘nested’ case-control, the review used the combined STROBE Statement for cohort, cross-sectional and case-control designs, and its extensions, to assess the reporting standards in UKB articles.

# Methods

# References

### 2021-08-18 manuscript/manuscript.Rmd
expand draft

@@
 knitr::opts_chunk$set(echo = TRUE)
+```
+
+# Abstract
+
+Background
+
+Objectives
+
+Data sources
+
+Eligibility criteria
+
+Participants
+
+Intervention
+
+Study appraisal
+
+Synthesis methods
+
+Restuls
+
+Limitations
+
+Conclusions
+
+Implications
+
+```{r, prisma}
+
+# 4. questions picos
+
 ```
@@
 
+We aimed to conduct a systematic review of articles reporting studies using data from the UKB with the aim of answering the following questions:
+
+1. What is the quality of reporting of findings from the UKB (in terms of the detail, clarity and completeness with which authors report their design, analysis and results)?
+2. Does reporting quality vary across the type of studies conducted?
+
 # Methods
 
+## Protocol registration
+
+This study was preregistered before we began to screen the search results [@drax2019].
+
+## Eligibility criteria
+
+Eligible articles are full research articles that report analyses of UKB data. Each article that is excluded at the full-text level of screening will be listed with the reason for exclusion.
+
+Inclusion criteria:
+* Publication date: After 30 March 2012. 
+This is the date the UKB was launched and UKB data was first available beforehand.
+* Publication type: Full research articles (including simulation and modelling articles).
+* Language: English.
+Relevant studies written in another language may be eligible but, to ensure accurate and precise evaluation, the article would require a professional translation which is not feasible given the review’s resources.
+* Data source: Article’s findings were produced from analysis(es) of data from the UKB. 
+This includes studies which reported findings from other studies alongside findings from UKB data.
+
+Exclusion criteria:
+* Article findings were produced from analysis(es) of pooled data from multiple studies.
+These are studies which have used data from participants from the UKB and other datasets in the same analysis.
+* Meta-analyses; narrative and systematic reviews; preprints; post-prints; replies; letters to the editor; corrections; and any other publication types that are not full research articles.
+* Retracted articles
+
+## Information sources and search strategy
+
+Four databases were electronically searched on 15/01/2019 for UKB articles: PubMed, EMBASE, Web of Science Core Collection (WoSCC) and PsycINFO. 
+The search strategy was limited to articles published from 2012 onwards and there was no restriction on language or publication type.
+
+The full search strategy for each database is included in Appendix A of the protocol [osf ref]. 
+To summarise, the terms "UK Biobank", "UKB", "UKBiobank", and "UKB Resource" were included in the "Title", "Abstract", "Keywords", and "All Fields" fields or equivalent of each database.
+
+```{r}
+# is there counts for the number of articles from th eUKB list?
+
+```
+
+## study selection
+
+We conducted four stages of screening for eligibility.
+First, using Endnote, KD removed duplicate results, checked the results for completeness, updated the metadata if necessary, and retrieved the full-texts of all results.
+KD checked completeness using a list of published full research articles that have used the UKB Resource, provided by the UKB on 27 November 2018. 
+Using the full-texts of the results, KD assessed if results were full research articles, and excluded reviews, corrections, conference abstracts, etc.
+
+Second, the results were imported into Rayyan [@ouzzani2016] and screened by BW and KD to determine whether they met the inclusion criteria.
+If insufficient information was available to determine a paper’s eligibility, the authors were contacted to request the required information. 
+If this information was unobtainable, the article was excluded.
+Although contacting the authors was our intention, time constraints meant we excluded some articles with insufficient information without contacting the authors.
+
+```{r}
+# check rayyan
+```
+
+In the protocol, we stated that data extraction would follow screening. 
+During which studies would be "classified according to their data type".
+This was so we could analyse articles according to their appropriate STROBE extension, such as the strengthening the reporting of genetic association studies (STREGA) statement [@little2009].
+Given that it would be too time-consuming to extract data from all studies in all included articles and to design multiple data extraction forms, we decided to extract data and assess the reporting quality of traditional epidemiology studies only.
+This means the third step was allocating articles to different classification groups according to the type of data they analysed.
+The classification criteria were refined after piloting them on 20 studies were:
+
+*	Traditional epidemiology – studies investigating associations between exposures and health outcomes using the UK Biobank data alone.
+
+*	Mendelian randomisation – studies that conduct Mendelian randomisation analyses on UK Biobank data alone
+
+*	Other – articles that do not meet the ‘Traditional epidemiology’ or ‘MR’ classification criteria described above or include analyses of genetic or imaging data.
+
+Articles abstracts were classified independently by MG and KD. 
+Full texts were examined where classification is unclear from the abstract.
+Any conflicts were discussed by KD and BW and resolved by mutual consent if possible. 
+If not resolved after discussion RR assign the category.
+Articles could be allocated into more than one classification group.
+
+### Further exclusions
+```{r}
+# add details of articles excluded after study design classification
+```
+KD examined the notices attached to the online version of the observational epidemiology articles
+to identify any retracted articles. 
+None were identified
+
+1 article was incorrectly classified as an ‘observational epidemiology’ study so 177 observational epidemiology articles remain.
+
+## Data collection process
+
+### Study Design Form
+
+Different STROBE items are relevant for cohort, case-control and cross-sectional study designs. 
+To ensure coders completed the same STROBE items for each article KD and MG independently identified the study designs used by each of the 178 “observational epidemiology” articles by completing the Study Design Form for each article.
+In the form, extractors also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
+Any conflicts were discussed by the two extractors and resolved by mutual consent if possible.
+If not resolved after discussion, RR determined the study designs. 
+The Study Design Form was hosted on
+Qualtrics but a PDF version can be found at https://o
+
+The articles were randomly assigned an ID number from 1-178. 
+The articles with an id number between 1
+and 80 were selected for the first round of data extraction. 
+KD was assigned as the first extractor to all 80 articles. 
+The second extractor was randomly assigned from MG, RR or BW.
+We initially randomly selected 80 articles with the hope that we would randomly select more studies once data extraction for the first 80 were complete.
+Unfortunately we did not have time to assess more studies.
+
+### Data Extraction Form
+
+The Data Extraction Form was hosted on Qualtrics and piloted twice. 
+This form combined the data items described as part of the "Data Extraction Form" and "Reporting Quality Assessment" in the protocol, but not all were kept.
+In the first, 5 observational epidemiology articles were piloted by MG, KD and/or RR. 
+Major changes were made so the Data Extraction Form will be completed again for the articles in the first pilot. 
+In the second, KD piloted the changed form on 3
+articles. 
+No changes were made so the responses for articles in the second pilot were retained.
+
+The extractors independently completed the Data Extraction Form for each article assigned to them.
+Initially, any conflicts were discussed by the two extractors and resolved by mutual consent if possible. 
+If not resolved after discussion, RR resolved the conflict.
+However, given the number of conflicts this could not be sustained.
+To reduce the time-burden on other authors KD and RR devised a series of "Rules" KD could use to resolve conflicts by herself. 
+This means the majority of conflicts were resolved by these "Rules" or KD's own judgement.
+
+### Other data collection
+
+The items in Table 1 were extracted manually by KD or automatically by Endnote which contained the article’s metadata exported from the database the article was located in.
+
+### Unclear and Missing Information.
+
+In our protocol we planned to locate and examine publications linked to the article, e.g.
+related full research articles, comments, corrections, etc, so that we could retrieve any missing items in the Data Extraction Form from them.
+If the missing information is not contained in a linked publication we planned to contact the article’s authors.
+However, we realised this would give an inaccurate assessment of the verision of record. 
+Also few articles had linked publications and that there was a lot of missing or unclear information.
+Therefore, we did not attempt to retrieve any missing items in the Data Extraction Form, from either the authors or linked articles.
+
+# Results
+
+## Study selection
+
+```{r}
+bw <-read.csv("../outputs//kd-bw-resolved.csv", stringsAsFactors = F, encoding = "UTF-8", na.strings = "")
+mg <-read.csv("../outputs/kd-mg-resolved.csv", stringsAsFactors = F, encoding = "UTF-8", na.strings = "")
+rr <-read.csv("../outputs/kd-rr-resolved.csv", stringsAsFactors = F, encoding = "UTF-8", na.strings = "")
+
+drop_cols <- function(df){
+  suf_drop <- c(".bw", ".rr", ".mg", ".correct")
+  suf_pat <- paste0("\\", suf_drop, "$", collapse ="|")
+  df <- df[, -grep(suf_pat, colnames(df))]
+  return(df)
+}
+
+bw <- drop_cols(bw)
+mg <- drop_cols(mg)
+rr <- drop_cols(rr)
+```
+
+## Study characteristics
+
+## Reporting quality
+
+# Discussion
+
+## Summary
+
+## Limitations
+
+## Conclusions
+
+# Funding Source
+
+This review is being funded by the John Climax Benevolent Fund. 
+The funder will support the conduct of the review by paying KD’s stipend. 
+The funder had no input on any aspect of the project, such the protocol design, data collection, data analysis nor interpretation or publication of results.
+
+# Conflicts of Interest
+There are no conflicts of interest to report.
+
 # References

### 2021-11-03 manuscript/manuscript.Rmd
add analysis to manuscript

@@
 ---
-title: "The UK Biobank Study: Systematic Review of the Reporting Quality in UK Biobank Studies"
-author: "Emma Anderson, George Davey Smith, Katie Drax, Mark Gibson, Marcus Munafò, Benjamin Woolf, and Rebecca Richmond"
+title: 'The UK Biobank Study: Systematic Review of the Reporting Quality in UK Biobank
+  Studies'
+author: Emma Anderson, George Davey Smith, Katie Drax, Mark Gibson, Marcus Munafò,
+  Benjamin Woolf, and Rebecca Richmond
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
   word_document:
-      number_sections: true
+    number_sections: yes
+  html_document:
+    df_print: paged
+  pdf_document: default
+csl: apa.csl
 bibliography: manuscript.bib
-csl: apa.csl
 ---
 
-```{r setup, include=FALSE}
-knitr::opts_chunk$set(echo = TRUE)
+```{r setup, include=FALSE, echo=F}
+knitr::opts_chunk$set(echo = F)
+
+############
+# libraries ####
+#############
+library(magrittr)
+library(ggplot2)
+library(tableone)
+library(gtools)
+library(tidyr)
+library(RColorBrewer)
+library(dplyr)
+library(irr)
+library(lpSolve)
+
+######################
+# import  ####
+######################
+
+# save naming convention used for strobe item variables as a pattern to be used in grep
+
+strobe_pattern <- "^s[[:digit:]]{1,2}"
+
+import_recoded <- function(){
+  # most variables are unordered factors (nominal with set responses) so read in with stringsAsFactors = T
+  df_orig <- read.csv("../outputs/data-extraction-form-clean.csv", encoding = "UTF-8", na.strings = c(""), stringsAsFactors = T)
+  df <- df_orig
+  for(i in colnames(df)){
+    # remove - from "Partially-External" as R doesn't like punctuation in factor levels
+    x <- gsub("Partially-External", "PartiallyExternal", df[[i]]) %>%
+      # make not applicable lowercase so doesn't clash with missing (NA) notation
+      gsub("NA", "na", .) %>%
+      # convert back into factor as gsub will have converted it to character
+      as.factor(.)
+    # check removed
+    if(length(which(x == "Partially-External") != 0)) stop("some Partially-External remain in ", i)
+    if(length(which(x =="NA") != 0)) stop("some not applicable values remain in ", i)
+    df[[i]] <- x
+  }
+  return(df)
+}
+
+df <- import_recoded()
+
+###################
+# add presence cols ####
+#####################
+
+# add non-strobe cols indicating if non-strobe items present
+
+add_pres <- function(df){
+  df$ukb_app_pres <- !is.na(df$ukb_app)
+  df$email_pres <- !is.na(df$email)
+  df$keywords_pres <- !is.na(df$keywords)
+  df$coi_pres <- !is.na(df$coi)
+  # drop cols just created presence cols for as no longer needed for frequencies
+  df <- df[, -which(colnames(df) %in% c("ukb_app", "email", "keywords", "coi"))]
+  return(df)
+}
+
+df <- add_pres(df)
+
+############
+# drop non-strobe cols ####
+#############
+
+drop_cols <- function(){
+  # remove predict col as df only contains non-prediction papers
+  if(!any(df$predict == "Yes", na.rm = T)){
+    df$predict <- NULL
+  }else{
+    stop("predict col not empty")
+  }
+  # drop access_article as this should be all yes
+  if(all(df$access_article == "Yes")){
+    df$access_article <- NULL
+  } else {
+    stop("access_article not just yes")
+  }
+  # save names of cols not needed 
+  void <- grep("title|comments\\.", colnames(df), value = T)
+  if(!all(void %in% colnames(df))) stop("df doesn't have ", void[!void %in% colnames(df)])
+  # drop not needed cols by name
+  df <- df[, -which(colnames(df) %in% void)]
+}
+
+df <- drop_cols()
+
+# vector of original column names for checking later
+
+if(sum(duplicated(colnames(df))) == 0){
+  df_cols <- colnames(df)
+} else {
+  stop("duplicated col names")
+}
+
+########################
+# save vector of strobe cols ####
+########################
+
+strobe_cols_names <- grep(strobe_pattern, colnames(df), value =T)
+
+strobe_stem <- gsub("_.*", "", strobe_cols_names) %>%
+  unique()
+
+###################################
+# split into strobe and non-strobe ####
+#################################
+
+# create funciton to check strobe items named & positioned correctly
+check_strobe <- function(df){
+  # find all strobe cols
+  strobe_cols_positions <- which(colnames(df) %in% strobe_cols_names)
+  # vector of numbers in strobe col names
+  strobe_nums <- strobe_cols_names %>%
+    # remove all punctuation and letters
+    gsub("[[:punct:]]|[[:alpha:]]", "", .)%>%
+    # remove duplicates
+    unique()
+
+  # check all strobe_nums are in 1-22
+  if(!identical(unique(gsub("[[:alpha:]]", "", strobe_nums)), as.character(c(1:22)))){
+    df <- NULL
+    warning("strobe_nums don't equal 1-22")
+  } 
+  if(sum(duplicated(strobe_nums)) != 0){
+    df <- NULL
+    warning("duplicate strobe_nums")
+  } 
+  # check positions of strobe items are sequential (i.e. all positions have a diff of 1 between them)
+  if(!all(abs(diff(strobe_cols_positions)) == 1)){
+    df <- NULL
+    warning("strobe items are not in sequential in df")
+  } 
+  return(df)
+}
+
+df <- check_strobe(df)
+
+# if passes strobe check, separate df into strobe items and non-strobe items
+
+s_df <- df[, which(colnames(df) %in% c("article_id", strobe_cols_names))]
+not_s_df <- df[, -which(colnames(df) %in% strobe_cols_names)]
+
+# check total ncol of s_df and not_s_df is same as original df +1 because article_id col is in both
+if(ncol(s_df) + ncol(not_s_df) != ncol(df) +1) stop("wrong number of cols in strobe or not_strobe dfs")
+
+###################
+# frequencies non-strobe ####
+##################
+
+# save version of not strobe df without qualtrics metadata variables to create table 1
+
+create_table1_df <- function(){
+  # save pattern to find qualtrics variables
+  qual <- c("distribution_channel", "duration_in_seconds", "end_date", "finished", "progress","recorded_date", "response_id", "response_type", "start_date", "user_language")
+  qual_pat <- paste0("^", qual, ".", collapse = "|")
+
+  # find qualtrics columns in df
+  qual_cols <- grep(qual_pat, colnames(not_s_df))
+
+  # check there are 4 of each qual column
+  if(length(qual_cols) != 4* length(qual)) stop("some qualtrics cols missing")
+
+  table1 <- not_s_df[, -qual_cols]
+
+  # drop article_id col as don't need freqs of these
+  table1 <- table1[, -which(colnames(table1) %in% "article_id")]
+
+  #convert all non-strobe cols into characters as easier than setting multiple different factor levels
+  for(i in colnames(table1)){
+    table1[[i]] <- as.character(table1[[i]])
+  }
+
+  # ensure UK and USA country values are capitalised as frequency is case-sensitive
+  uk_usa <- grep("^uk$|^usa$", table1$country, ignore.case = T)
+  table1$country[uk_usa] <- toupper(table1$country[uk_usa])
+  return(table1)
+}
+
+table1 <- create_table1_df()
+
+CreateTableOne(data = table1, includeNA = F)%>%
+  print(., noSpaces = T, showAllLevels = T) %>%
+  # export
+  write.csv(., "../outputs/table1_all_levels.csv")
+
+CreateTableOne(data = table1, includeNA =F )%>%
+  print(., noSpaces = T) %>%
+  # export
+  write.csv(., "../outputs/table1_no_levels.csv")
+
+#########################
+# recode strobe factors ####
+########################
+
+# recode strobe variables into factors so can analyse later
+factorise_strobe <- function(){
+  # check strobe cols correct
+  if(!all(strobe_cols_names %in% colnames(s_df))) stop("some strobe cols not in s_df")
+  # save s_df before recoding factors
+  pre_recode <- s_df
+  # recode all strobe cols in s_df
+  for(i in strobe_cols_names){
+    # set all cols to factor
+    s_df[[i]] <- as.factor(s_df[[i]])
+    #recode all factor levels
+    levels(s_df[[i]]) <- list(Yes = "Yes", PartiallyExternal = "PartiallyExternal", Partially = "Partially", No = "No", Unsure = "Unsure", na = "na")
+  }
+
+  # check recode successful
+
+  x <- c()
+
+  for(i in strobe_cols_names){
+    y <- all.equal(as.character(s_df[[i]]), as.character(pre_recode[[i]]))
+    x <- c(x, y)
+  }
+  # check all cols identical as they were before recoding
+  if(sum(x) != length(x) | length(x) == 0) stop("factor recoded cols in s_df are different to before recoding")
+  # check factors created
+  classes <- sapply(s_df, class)
+  if(!any(classes == "factor")) stop("no factors created")
+  return(s_df)
+}
+
+s_df <- factorise_strobe()
+
+#######################
+# divide strobe into ternary ####
+########################
+
+recode_strobe <- function(style = c("binary", "ternary")){
+  # check style arg correct
+  if(!style %in% c("binary", "ternary")) stop("style arg incorrect")
+
+  # recode s_df
+  x <- s_df
+
+  # convert all cols into character variables
+  x[] <- lapply(x, as.character)
+
+  # get ids of all strobe cols
+  id <- grep(strobe_pattern, colnames(x))
+
+  # check all are characters
+  for(i in id){
+    if(is.character(x[, i]) == F) stop(i, "isn't a character")
+  }
+  # recode into binary
+  if(style == "binary"){
+    for(i in id){
+      x[, i] <-  gsub("^Yes$", "1", x[, i]) %>%
+        gsub("^No$|^Partially$|^Unsure$|^PartiallyExternal$", "0", .) %>%
+        gsub("na", NA, .) %>%
+        as.numeric(.)
+      if(is.numeric(x[, i]) == F) stop(i, "not numeric")
+    }
+  }
+  # recode into ternary
+  if(style == "ternary") {
+    for(i in id){
+      x[, i] <-  gsub("^Yes$|^PartiallyExternal$", "1", x[, i]) %>%
+        gsub("^Partially$", "0.5", .) %>%
+        gsub("^No$|^Unsure$", "0", .) %>%
+        gsub("na", NA, .) %>%
+        as.numeric(.)
+      if(is.numeric(x[, i]) == F) stop(i, "not numeric")
+    }
+  }
+  #check strobe cols
+  if(!all(strobe_cols_names %in% colnames(x))) stop("some strobe cols not in s_df_tri")
+
+  return(x)
+}
+
+s_df_bin <- recode_strobe(style = "binary")
+s_df_tri <- recode_strobe(style = "ternary")
+
+#######################
+# calculate strobe scores for ternary  ####
+########################
+
+sum_items <- function(df, style = c("binary", "ternary")){
+  # check style arg correct
+  if(!style %in% c("binary", "ternary")) stop("style arg incorrect")
+
+  pre <- ncol(df)
+  for(i in strobe_stem){
+    # save dataframe of variables for same strobe item so can sum rows
+    x <- df[,grep(i, colnames(df))]
+    # create sum col name
+    sum_name <- paste(i, "sum", sep = "_")
+    # sum rows in x if it is a dataframe 
+    if(!is.numeric(x)){
+      sum_var <- rowSums(x, na.rm = T) / rowSums(!is.na(x))
+    } else {
+      # if not dataframe then no need to sum rows because only 1 col
+      sum_var <- x
+    }
+    # check no rows dropped during summing
+    if(length(sum_var) != nrow(df)) stop("missing some assessments")
+    # make scores binary or ternary 
+    if(style == "binary"){
+      # force into binary by rounding down any less than 1 to indicate all yes/not yes
+      sum_var[sum_var < 1] <- 0
+    }
+    if(style == "ternary"){
+      # force into ternary by rounding any not 0 or 1 numbers to 0.5 to indicate partialness
+      sum_var[sum_var >0 & sum_var < 1] <- 0.5
+    }
+    # add sum to df under sum_name
+    df[[sum_name]] <- sum_var
+  }
+
+  if(ncol(df) != pre + length(unique(strobe_stem)))
+    stop("cols don't equal number of n cols before function plus number strobe items")
+
+  # drop all columns that are not summs composite strobe items and strobe items without subdivisions
+  df <- df[, which(colnames(df) %in% c("article_id", grep("_sum$", colnames(df), value = T)))]
+
+  # remove NaN values because rowSums will have introduced them
+  is.nan.data.frame <- function(x)
+    do.call(cbind, lapply(x, is.nan))
+
+  df[is.nan(df)] <- NA
+
+  # remove _sum from all colnames so matches strobe stems
+  colnames(df) <- gsub("_.*", "", colnames(df))
+
+  return(df)
+}
+
+tri_sum <- sum_items(s_df_tri, style = "ternary")
+bin_sum <- sum_items(s_df_bin, style = "binary")
+
+#####################
+# create bar chart data ###
+#######################
+
+create_bar_data <- function(df){
+  bar_data <- df[, which(colnames(df) %in% strobe_stem)]
+  if(ncol(bar_data) != length(strobe_stem)) stop("some strobe items missing")
+  # clean col names so will match names of strobe items in strobe guidance
+  colnames(bar_data) <- colnames(bar_data) %>%
+    gsub("^s|__sum|starred", "", .)
+
+  # change all to character to create table one
+  bar_data[] <- lapply(bar_data, as.character)
+
+  bar_data <- tidyr::pivot_longer(bar_data, colnames(bar_data), names_to = "strobe_item", values_to = "value") 
+  x <- table(bar_data$strobe_item, bar_data$value, useNA = "always")
+  y <- prop.table(table(bar_data$strobe_item, bar_data$value), 1) %>%
+    round(., 2)
+  # convert to dataframe
+  x <- as.data.frame.matrix(x)
+  y <- as.data.frame.matrix(y)
+
+  # row names will contain the x values given to table(x, y), move these to a column >
+  # remove the X added
+  x$strobe_item <- gsub("^X", "", row.names(x))
+  row.names(x) <- NULL
+
+  y$strobe_item <- gsub("^X", "", row.names(y))
+  row.names(y) <- NULL
+
+  # useNA in x means there will be an blank column name for the NA values, rename this as won't pivot if any missing
+  colnames(x)[is.na(colnames(x))] <- "na"
+  # useNA also creates an "NA" row, remove this
+  x <- x[-grep("NA", x$strobe_item), ]
+
+  # save response cols so can pivot them
+  x_cols <- colnames(x)[-which(colnames(x) %in% "strobe_item")]
+
+  # create response column
+  x <- pivot_longer(x, all_of(x_cols), names_to = "response", values_to = "count")
+
+  # save response cols so can pivot them
+  y_cols <- colnames(y)[-which(colnames(y) %in% "strobe_item")]
+
+  # create response column
+  y <- pivot_longer(y, all_of(y_cols), names_to = "response", values_to = "percent")
+
+  # check strobe items identical so will merge properly
+  if(!identical(unique(x$strobe_item), unique(y$strobe_item))) stop("x and y strobe items different")
+
+  bar_data <- full_join(x, y, by = c("strobe_item", "response"), suffix = c(".count", ".percent"))
+
+  # reorder
+  bar_data <- bar_data[mixedorder(as.character(bar_data$strobe_item)),]
+
+  return(bar_data)
+}
+
+bar_data <- create_bar_data(tri_sum)
+
+#############################
+# export data to create bar chart labels ####
+##########################
+
+create_bar_labels_csv <- function(){
+  strobe_qs <- read.csv("../outputs/extraction_dict.csv", stringsAsFactors = F, encoding = "UTF-8") %>%
+    .[, colnames(.) %in% c("question", "variable")] %>%
+    # select those that are strobe items 
+    .[grep("s[[:digit:]]{1,2}", .$variable), ] %>%
+    .[-grep("\\_ev|\\_star", .$variable), ]
+
+  strobe_qs$question <- sub("[^0-9]*", "", strobe_qs$question)
+  strobe_qs$strobe_item <- sub("\\_.*", "", strobe_qs$variable)
+  write.csv(strobe_qs, "../outputs/bar_labels.csv", row.names = F, fileEncoding = "UTF-8")
+}
+
+create_bar_labels_csv()
+
+##############
+## MANUAL ####
+############
+
+# MANUALLY INSTRUCTIONS
+# added text describing the strobe item the bar_labels.csv
+
+###########################
+# import labels and clean ####
+########################
+
+create_labels <- function(){
+
+  # import general as these contain labels at the strobe item level (not design specific)
+  labels <- read.csv("../data/bar_labels.csv", encoding = "UTF-8", stringsAsFactors = F, na.strings = "") %>%
+    .[, colnames(.) %in% c("variable", "item_level_label")]
+
+  # clean variable names so matches bar_chart_freq colnames (i.e. no subdivisions and no design specific questions) >
+  # easiest to do this by cleaning non-cohort specific quesitons first then cohort excluding 14c_coh since this contains no roman numerials and is not a sub division
+  labels$variable <- gsub("_.*|starred|s", "", labels$variable)
+
+  # remove duplicated rows now variables cleaned
+  labels <- labels[!duplicated(labels),]
+  # sort
+  labels <- labels[mixedorder(as.character(labels$variable)),]
+
+  # select applic col so can merge in
+  applic_df <- bar_data[bar_data$response == "na", which(colnames(bar_data) %in% c("strobe_item", "count"))]
+
+  # add applic col to labels rows only so can remove empty labels that are empty and wont appear in bar chart
+  labels <- left_join(labels, applic_df, by = c("variable" = "strobe_item"))
+
+  # add number of applicable items to labels so can see how many items were relevant in bar
+  if(identical(labels$variable, 
+               as.character(unique(bar_data$strobe_item)))){
+    # paste variable together with description , separate by space so can gsub later
+    labels$x_labels <- paste(labels$variable, labels$item_level_label, sep = " ") %>%
+      # order
+      .[gtools::mixedorder(.)]%>%
+      # wrap
+      stringr::str_wrap(., width = 55) %>%
+      # add n = applicable on a newline
+      paste(., " (n=", nrow(df) - labels$count, ")", sep = "")
+  } else {
+    stop("not identical")
+  }
+  return(labels)
+}
+
+labels <- create_labels()
+
+#######################
+## bar chart for all ####
+##################
+strobe_levels <-  bar_data$strobe_item[!duplicated(bar_data$strobe_item)]
+# set levels as mixed order so order preserved in ggplot
+bar_data$strobe_item<- factor(bar_data$strobe_item, levels = 
+                                strobe_levels[gtools::mixedorder(strobe_levels)])
+
+bar_data$response <- as.factor(bar_data$response)
+levels(bar_data$response) <- c("No", "Partially", "Yes", "Not Applicable")
+
+# check labels identical to strobe item numbers in labels$x_labels (gsub at the space) to ensure labels will map to chart
+if(!identical(gsub(" .*", "", labels$x_labels), as.character(strobe_levels))) stop("labels won't map")
+
+png("Rplot9.png", width = 1300, height = 1500)
+
+# strobe item on x axis and plot % yes
+bar_data %>%ggplot(aes(x=strobe_item, y=count, fill = response)) + 
+  # blue fill
+  geom_bar(stat="identity") +
+  theme_classic() +
+  theme(text = element_text(size=30))+
+  scale_x_discrete(labels= labels$x_labels) +
+  scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
+  coord_flip() +
+  xlab("STROBE Item") +
+  ylab("Count") +
+  ggtitle("Figure 1. STROBE completion")+
+  scale_fill_brewer(palette = "Spectral") +
+  #geom_text(aes(label = count), vjust = 0, size = 8) +
+  theme(plot.title = element_text(size = 22)) 
+dev.off()
+
+#####################
+# calculate irr for NA and applicable ####
+################################
+
+inter <- read.csv("../outputs/data-extraction-form-conflicted.csv", na.strings = "")
+
+create_kappa_df <- function(df){
+
+  # save stems of variables that were not double coded (and therefore not conflicted)
+  # kd alone extracted some variables so save all these variables single/unique variables
+  single <- c("email","country","ukb_app","keywords","coi", "article_id") %>%
+    #  qualtrics metadata, comments & title substring vars are unique to each coder so stems duplicated but not colnames >
+    c(., "distribution_channel", "duration_in_seconds", "end_date", "finished", "progress", "recorded_date", "response_id", "response_type", "start_date", "user_language") %>%
+    # remove those unique to each coder
+    c(., "comments", "title_clean", "title_sub", "title")
+
+  stem <- grep("\\.", colnames(df), value = T) %>%
+    gsub("\\..*", "", .) %>%
+    unique() %>%
+    # remove single coded items since these can't be conflicted
+    .[!. %in% single] %>%
+    # remove predict because we already resolved these conflicts
+    .[!. %in% "predict"] %>%
+    # remove other variables that are not discrete
+    .[!. %in% c("other_guidelines", "reg_id")] %>%
+    # add '.' onto end so will match full word
+    paste0(., "\\.")
+
+  kappa <- c()
+  z <- c()
+  p_value <- c()
+  variable <- c()
+
+  for(i in stem){
+    x <- df[, grep(i, colnames(df))]
+    y <- kappa2(x, weight = "unweighted")
+    variable <- c(variable, i)
+    kappa <- c(kappa, y$value)
+    z <- c(z, y$statistic)
+    p_value <- c(p_value, y$p.value)
+  }
+
+  irr_df <- data.frame(variable = variable, kappa= kappa, z= z, p_value = p_value)
+
+}
+
+# divide df into into NA and applicable
+
+divide_na <- function(){
+  for(i in colnames(inter)){
+    x <- inter[[i]]
+    x[x != "NA"] <- "A"
+    inter[[i]] <- x
+  }
+  return(inter)
+}
+
+inter_na <- divide_na()
+na_kappa <- create_kappa_df(inter_na)
+
+#######################
+# calculate strobe scores for ternary for inter  ####
+########################
+
+# split into kd and coder2 so can sum strobe items for each then rejoin
+kd <- inter[, c("article_id", grep("\\.kd", colnames(inter), value = T))]
+coder2 <- inter[, c("article_id", grep("\\.coder2", colnames(inter), value = T))]
+
+sum_items <- function(df, coder_suffix){
+  # set "NA" and "Unsure" as missing so calculations only apply to reporting judgements
+  df[df == "NA" | df == "Unsure"] <- NA
+  strobe_cols_names <- grep("^s[[:digit:]]{1,2}", colnames(df), value =T)
+
+  # remove everything after _ or suffix
+  strobe_stem <- gsub("_.*|\\..*", "", strobe_cols_names) %>%
+    unique()
+
+  pre <- ncol(df)
+
+  for(i in strobe_stem){
+
+    # save dataframe of variables for same strobe item so can sum rows
+    x <- df[,grep(i, colnames(df))]
+    # create sum col name
+    sum_name <- paste0(i, "_sum", coder_suffix)
+    # create empty sum_var to add values to in for loop
+    sum_var <- c()
+    # sum rows in x if it is a dataframe 
+    if(!is.character(x)){
+      for(j in 1:nrow(x)){
+        if(all(x[j, ] %in% c("Yes", "Partially-External"), na.rm = T)) {
+          sum_var[j] <- "Yes"
+        } else {
+          if(!all(x[j, ] %in% c("Yes", "Partially-External"), na.rm = T) & !all(x[j, ] == "No", na.rm = T)) {
+            sum_var[j] <- "Partially"
+          } else {
+            if(all(x[j, ] == "No", na.rm = T)) {
+              sum_var[j] <- "No"
+            } else {
+              stop(j, " in ", i, " doesn't contain just yes, pe, partially or no")
+            }
+          }
+        }
+      }
+    } else {
+      # if not dataframe then no need to sum rows because only 1 col
+      sum_var <- x
+    }
+    # check no rows dropped during summing
+    if(length(sum_var) != nrow(df)) stop("missing some assessments")
+    # add sum to df under sum_name
+    df[[sum_name]] <- sum_var
+  }
+  if(ncol(df) != pre + length(unique(strobe_stem)))
+    stop("cols don't equal number of n cols before function plus number strobe items")
+
+  # drop all cols that are not strobe sums & article id so can merge on id
+  df <- df[, which(colnames(df) %in% c("article_id", grep("_sum", colnames(df), value =T)))]
+
+  return(df)
+}
+
+kd <- sum_items(kd, ".kd")
+coder2 <- sum_items(coder2, ".coder2")
+
+inter_item <- full_join(kd, coder2, by = "article_id") %>%
+  # reorder
+  .[ , gtools::mixedorder(colnames(.))] %>%
+  # remove article_id
+  .[, -which(colnames(.) %in% "article_id")]
+
+kappa_item <- create_kappa_df(inter_item)
+
 ```
@@
 Inclusion criteria:
+
 * Publication date: After 30 March 2012. 
@@
 If this information was unobtainable, the article was excluded.
-Although contacting the authors was our intention, time constraints meant we excluded some articles with insufficient information without contacting the authors.
+Time constraints meant we excluded some articles with insufficient information without contacting the authors.
 
-```{r}
-# check rayyan
-```
-
-In the protocol, we stated that data extraction would follow screening. 
-During which studies would be "classified according to their data type".
-This was so we could analyse articles according to their appropriate STROBE extension, such as the strengthening the reporting of genetic association studies (STREGA) statement [@little2009].
-Given that it would be too time-consuming to extract data from all studies in all included articles and to design multiple data extraction forms, we decided to extract data and assess the reporting quality of traditional epidemiology studies only.
-This means the third step was allocating articles to different classification groups according to the type of data they analysed.
-The classification criteria were refined after piloting them on 20 studies were:
+Third, We classified studies as "traditional epidemiology", "Mendelian randomisation" and "Other", according to the type of data they analysed.
+After being piloted on 20 studies, the classification criteria were:
 
 *	Traditional epidemiology – studies investigating associations between exposures and health outcomes using the UK Biobank data alone. 
-
 *	Mendelian randomisation – studies that conduct Mendelian randomisation analyses on UK Biobank data alone
-
 *	Other – articles that do not meet the ‘Traditional epidemiology’ or ‘MR’ classification criteria described above or include analyses of genetic or imaging data.
@@
 Articles could be allocated into more than one classification group.
+We then excluded all but the traditional epidemiology studies.
+The Mendialian randomisation studies were not relevant to this paper but we included it as a classification because the data may be used for one of our author's future projects.
 
-### Further exclusions
-```{r}
-# add details of articles excluded after study design classification
-```
-KD examined the notices attached to the online version of the observational epidemiology articles
-to identify any retracted articles. 
-None were identified
+Only including observational epidemiological studies is a major departure from the protocol in which we stated that studies would be "classified according to their data type" during data extraction but that all classifications would be included.
+This was so we could analyse articles according to their appropriate STROBE extension, such as the strengthening the reporting of genetic association studies (STREGA) statement [@little2009].
+Given that it would be too time-consuming to extract data from all studies in all included articles and to design multiple data extraction forms, we decided to extract data and assess the reporting quality of studies that could be assessed using the original STROBE guidelines only, i.e.
+traditional epidemiology studies.
 
-1 article was incorrectly classified as an ‘observational epidemiology’ study so 177 observational epidemiology articles remain.
+Fourth, we catergorised the 178 studies classified as observational epidemiological studies into those with cohort, case-control and cross-sectional study designs using the Study Design Form described below.
+During this process KD also examined the notices attached to the online version of the observational epidemiology articles to identify any retracted articles. 
+None were identified.
+However, we did identify studies that should be excluded, these studies were:
+
+* Studies incorrectly classified as observational epidemiological studies.
+* Studies with inaccessible supplementary material
+* Studies we did consider to be observiational epidemiological studies but could not consider them to have cohort, case-control or cross-sectional designs - for example those using prediction models
 
@@
 If not resolved after discussion, RR determined the study designs. 
-The Study Design Form was hosted on
-Qualtrics but a PDF version can be found at https://o
+The Study Design Form was hosted on Qualtrics but a PDF version can be found at https://osf.io/jfk24/.
 
 The articles were randomly assigned an ID number from 1-178. 
-The articles with an id number between 1
-and 80 were selected for the first round of data extraction. 
+The articles with an ID number between 1 and 80 were selected for the first round of data extraction. 
 KD was assigned as the first extractor to all 80 articles. 
@@
 Major changes were made so the Data Extraction Form will be completed again for the articles in the first pilot. 
-In the second, KD piloted the changed form on 3
-articles. 
+In the second, KD piloted the changed form on 3 articles. 
 No changes were made so the responses for articles in the second pilot were retained.
@@
 However, given the number of conflicts this could not be sustained.
-To reduce the time-burden on other authors KD and RR devised a series of "Rules" KD could use to resolve conflicts by herself. 
+To reduce the time-burden on other authors KD and RR devised a series of "rules" and general guidelines KD could use to resolve conflicts by herself. 
 This means the majority of conflicts were resolved by these "Rules" or KD's own judgement.
+An explanation of the rules and guidelines can be found at https://osf.io/hngcj/.
 
@@
 
-```{r}
-bw <-read.csv("../outputs//kd-bw-resolved.csv", stringsAsFactors = F, encoding = "UTF-8", na.strings = "")
-mg <-read.csv("../outputs/kd-mg-resolved.csv", stringsAsFactors = F, encoding = "UTF-8", na.strings = "")
-rr <-read.csv("../outputs/kd-rr-resolved.csv", stringsAsFactors = F, encoding = "UTF-8", na.strings = "")
-
-drop_cols <- function(df){
-  suf_drop <- c(".bw", ".rr", ".mg", ".correct")
-  suf_pat <- paste0("\\", suf_drop, "$", collapse ="|")
-  df <- df[, -grep(suf_pat, colnames(df))]
-  return(df)
-}
-
-bw <- drop_cols(bw)
-mg <- drop_cols(mg)
-rr <- drop_cols(rr)
-```
-
 ## Study characteristics
@@
 ## Reporting quality
+
+![Figure 1](Rplot9.png)
 

### 2021-11-03 rscripts/03-analysis/01-analysis.R

@@ -1,642 +1,0 @@
-############
-# libraries ####
-#############
-library(magrittr)
-library(ggplot2)
-library(tableone)
-library(gtools)
-library(tidyr)
-library(RColorBrewer)
-library(dplyr)
-library(irr)
-library(lpSolve)
-
-######################
-# import  ####
-######################
-
-
-# save naming convention used for strobe item variables as a pattern to be used in grep
-
-strobe_pattern <- "^s[[:digit:]]{1,2}"
-
-import_recoded <- function(){
-  # most variables are unordered factors (nominal with set responses) so read in with stringsAsFactors = T
-  df_orig <- read.csv("outputs/data-extraction-form-clean.csv", encoding = "UTF-8", na.strings = c(""), stringsAsFactors = T)
-  df <- df_orig
-  for(i in colnames(df)){
-    # remove - from "Partially-External" as R doesn't like punctuation in factor levels
-    x <- gsub("Partially-External", "PartiallyExternal", df[[i]]) %>%
-      # make not applicable lowercase so doesn't clash with missing (NA) notation
-      gsub("NA", "na", .) %>%
-      # convert back into factor as gsub will have converted it to character
-      as.factor(.)
-    # check removed
-    if(length(which(x == "Partially-External") != 0)) stop("some Partially-External remain in ", i)
-    if(length(which(x =="NA") != 0)) stop("some not applicable values remain in ", i)
-    df[[i]] <- x
-  }
-  return(df)
-}
-
-df <- import_recoded()
-
-###################
-# add presence cols ####
-#####################
-
-# add non-strobe cols indicating if non-strobe items present
-
-add_pres <- function(df){
-  df$ukb_app_pres <- !is.na(df$ukb_app)
-  df$email_pres <- !is.na(df$email)
-  df$keywords_pres <- !is.na(df$keywords)
-  df$coi_pres <- !is.na(df$coi)
-  # drop cols just created presence cols for as no longer needed for frequencies
-  df <- df[, -which(colnames(df) %in% c("ukb_app", "email", "keywords", "coi"))]
-  return(df)
-}
-
-df <- add_pres(df)
-
-############
-# drop non-strobe cols ####
-#############
-
-drop_cols <- function(){
-  # remove predict col as df only contains non-prediction papers
-  if(!any(df$predict == "Yes", na.rm = T)){
-    df$predict <- NULL
-  }else{
[…]

### 2021-11-05 rscripts/02_data_extraction/03_merge_resolved_conflicts.R
set dup or NA cols to NA values don't drop, add in presence of SM col

@@ -278,11 +278,12 @@ create_inter_df <- function(){
     .[ , gtools::mixedorder(colnames(.))]
   # drop all correct cols because not necessary for irr calculation
   inter <- inter[, -grep(".correct$", colnames(inter))]
-  #
   # set all values for duplicates and not applicable items to NA
   drop_pat <- paste(c("5_vi", "s6a_iii", "10", "8starred_iii"), collapse = "|")
-  # drop na cols
-  inter <- inter[, -grep(drop_pat, colnames(inter))]
+  # set all to NA
+  for(i in grep(drop_pat, colnames(inter), value = T)){
+    inter[[i]] <- rep("NA", length(inter[[i]]))
+  }
   
   return(inter)
 }
@@ -438,13 +439,24 @@ df <- clean_values()
 # merge in epi_access ####
 #####################
 
+# add in whether they had supplementary material, this is in the first_80 ref csv
+first_80 <- read.csv("https://osf.io/9w72e/?action=download", encoding = "UTF-8", na.strings = c("", " ")) %>%
+  .[, which(colnames(.) %in% c("article_id", "access_supp"))]
+
+# I removed Davie from the first_80 in the first revision >
+# add this back in because we decided it did actually contain a cohort design
+
+first_80 <- rbind(first_80, c("Davie2018bank79-y", "Yes"))
+
 epi_access <- read.csv("data/epi_access.csv", stringsAsFactors = F, encoding = "UTF-8", na.strings = c("", " ")) %>%
   .[, which(!colnames(.) %in% c("doi", "title", "authors"))]
+
+epi_access <- left_join(first_80, epi_access, by = "article_id")
 
 df <- left_join(df, epi_access, by = "article_id")
 
 # TEMPORARY REPLACE BECKY IT VALUES AS NA
-
+ 
 for(i in colnames(df)){
   if(any(grepl("becky", df[[i]]))) print(i)
   becky_its <- grep("becky it", df[[i]], ignore.case = T)

### 2021-11-05 manuscript/manuscript.Rmd
replace table1 with table2, don't recode unsure because should all be resolved, reorder bar labels, format kappa tables and add all NA cols, create supp table 2, add narrative results

@@
 library(lpSolve)
+library(kableExtra)
+library(pander)
 
@@
   unique()
-
 ###################################
@@
 
-###################
-# frequencies non-strobe ####
-##################
-
-# save version of not strobe df without qualtrics metadata variables to create table 1
-
-create_table1_df <- function(){
-  # save pattern to find qualtrics variables
-  qual <- c("distribution_channel", "duration_in_seconds", "end_date", "finished", "progress","recorded_date", "response_id", "response_type", "start_date", "user_language")
-  qual_pat <- paste0("^", qual, ".", collapse = "|")
-
-  # find qualtrics columns in df
-  qual_cols <- grep(qual_pat, colnames(not_s_df))
-
-  # check there are 4 of each qual column
-  if(length(qual_cols) != 4* length(qual)) stop("some qualtrics cols missing")
-
-  table1 <- not_s_df[, -qual_cols]
-
-  # drop article_id col as don't need freqs of these
-  table1 <- table1[, -which(colnames(table1) %in% "article_id")]
-
-  #convert all non-strobe cols into characters as easier than setting multiple different factor levels
-  for(i in colnames(table1)){
-    table1[[i]] <- as.character(table1[[i]])
-  }
-
-  # ensure UK and USA country values are capitalised as frequency is case-sensitive
-  uk_usa <- grep("^uk$|^usa$", table1$country, ignore.case = T)
-  table1$country[uk_usa] <- toupper(table1$country[uk_usa])
-  return(table1)
-}
-
-table1 <- create_table1_df()
-
-CreateTableOne(data = table1, includeNA = F)%>%
-  print(., noSpaces = T, showAllLevels = T) %>%
-  # export
-  write.csv(., "../outputs/table1_all_levels.csv")
-
-CreateTableOne(data = table1, includeNA =F )%>%
-  print(., noSpaces = T) %>%
-  # export
-  write.csv(., "../outputs/table1_no_levels.csv")
-
 #########################
@@
     #recode all factor levels
-    levels(s_df[[i]]) <- list(Yes = "Yes", PartiallyExternal = "PartiallyExternal", Partially = "Partially", No = "No", Unsure = "Unsure", na = "na")
+    levels(s_df[[i]]) <- list(Yes = "Yes", PartiallyExternal = "PartiallyExternal", Partially = "Partially", No = "No", na = "na")
   }
@@
       x[, i] <-  gsub("^Yes$", "1", x[, i]) %>%
-        gsub("^No$|^Partially$|^Unsure$|^PartiallyExternal$", "0", .) %>%
+        gsub("^No$|^Partially$|^PartiallyExternal$", "0", .) %>%
         gsub("na", NA, .) %>%
@@
         gsub("^Partially$", "0.5", .) %>%
-        gsub("^No$|^Unsure$", "0", .) %>%
+        gsub("^No$", "0", .) %>%
         gsub("na", NA, .) %>%
@@
   bar_data <- tidyr::pivot_longer(bar_data, colnames(bar_data), names_to = "strobe_item", values_to = "value") 
-  x <- table(bar_data$strobe_item, bar_data$value, useNA = "always")
+  x <- table(bar_data$strobe_item, bar_data$value, useNA = "always") %>%
+    # convert to dataframe
+    as.data.frame.matrix()
   y <- prop.table(table(bar_data$strobe_item, bar_data$value), 1) %>%
-    round(., 2)
-  # convert to dataframe
-  x <- as.data.frame.matrix(x)
-  y <- as.data.frame.matrix(y)
+    # round so more interpretable
+    round(., 2) %>%
+    as.data.frame.matrix()
 
@@
 bar_data$response <- as.factor(bar_data$response)
-levels(bar_data$response) <- c("No", "Partially", "Yes", "Not Applicable")
+
+# set levels using list so is accurate
+levels(bar_data$response) <- list(NotApplicable = "na", No = 0, Partially = 0.5, Yes = 1)
+
+# add space to NA level since can't do this in list
+levels(bar_data$response)[levels(bar_data$response) == "NotApplicable"] <- "Not Applicable"
 
@@
   ggtitle("Figure 1. STROBE completion")+
-  scale_fill_brewer(palette = "Spectral") +
+  scale_fill_brewer(palette = "Greys") +
   #geom_text(aes(label = count), vjust = 0, size = 8) +
@@
 
+  # save stems of duplicate and not applicable cols so can join back in
+  na_items <- c("5_vi", "6a_iii", "10", "8starred_iii")
+  na_stem <-  grep(paste0(na_items, collapse ="|"), stem, value =T)
+
+  # remove all stems that are duplicate or not applicable as will throw error in kappa2 if set "NA" to missing and all are set to NA after qualtrics
+
+  stem <- stem[-which(stem == na_stem)]
+  #save empty objects so can save outputs from for loop
   kappa <- c()
@@
   variable <- c()
+  n_match <- c()
+  n_conflicts <-c()
+  n_na <- c()
 
@@
     p_value <- c(p_value, y$p.value)
+    n_match <- c(n_match, sum(x[[1]] == x[[2]], na.rm = T))
+    n_conflicts <- c(n_conflicts, sum(x[[1]] != x[[2]], na.rm = T))
+    # count number of items with NA--NA dyads or notNA--NA dyads so are excluded from kappa
+    n_na <- c(n_na, sum(is.na(x[[1]] != x[[2]])))
   }
 
-  irr_df <- data.frame(variable = variable, kappa= kappa, z= z, p_value = p_value)
+  # bind objects into df so can render in manuscript
+  irr_df <- data.frame(STROBEItem = variable, Kappa = kappa, Zstatistic = z, pvalue = p_value, Matches = n_match, Conflicts = n_conflicts, NotApplicableUnsure = n_na)
 
+  # create df of empty varibles and bind to main df so will appear in table
+  empty <- data.frame(STROBEItem = na_stem)
+  irr_df <- full_join(irr_df, empty, by = "STROBEItem")
+
+  # round all numeric cols so more interpretable
+  nums <- colnames(irr_df)[colnames(irr_df) != "STROBEItem"]
+  for(i in nums){
+    x <- irr_df[[i]]
+    if(!is.numeric(x)) stop(i, " col not numeric")
+    x <- round(x, digits = 2)
+    irr_df[[i]] <- x
+  }
+
+  # remove backslash that was added to variable names
+  irr_df$STROBEItem <- gsub("\\\\.*", "", irr_df$STROBEItem)
+
+  # reorder
+  irr_df <- irr_df[gtools::mixedorder(irr_df$STROBEItem), ]
+
+  # set rownames to null to ensure none created
+  row.names(irr_df) <- NULL
+
+  return(irr_df)
 }
@@
 inter_na <- divide_na()
-na_kappa <- create_kappa_df(inter_na)
+kappa_na <- create_kappa_df(inter_na)
 
@@
 kappa_item <- create_kappa_df(inter_item)
+
+kappa_item$STROBEItem <- gsub("^s|_.*|starred", "", kappa_item$STROBEItem)
 
@@
 Exclusion criteria:
+
 * Article findings were produced from analysis(es) of pooled data from multiple studies.
@@
 
-## Reporting quality
+Of the `r nrow(df)` included studies `r sum(grepl("Case-control", df$designs))` contained case-control designs, `r sum(grepl("Cohort", df$designs))` cohort, and `r sum(grepl("Cross-sectional", df$designs))` cross-sectional.
+These numbers add up to more than 70 because some articles contained cross-sectional and cohort studies.
+UK based researchers led the vast majority with `r sum(grepl("UK", df$country, ignore.case = F))` having authors affiliated with institutions in the UK.
+`r sum(grepl("2014", df$year))` studies were published in 2014, `r sum(grepl("2015", df$year))` in 2015, `r sum(grepl("2016", df$year))` in 2016, `r sum(grepl("2017", df$year))` in 2017, `r sum(grepl("2018", df$year))` in 2018, and `r sum(grepl("2019", df$year))` in 2019.
 
-![Figure 1](Rplot9.png)
+## STROBE Reporting quality
 
+Using the data extraction form on Qualtrics, we assessed reporting quality separately for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it.
+For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor and the applicability of the items was often uncertain. 
+We then summed the subdivisions to indicate the reporting quality of each STROBE item, these results are presented in Figure 1. 
+For completeness, the results for the subdivisions are available in Supplementary Table 2.
+
+We originally assessed reporting quality separately for all subdivisions of the STROBE items.
+The results for the subdivisions are reported in Supplementary Table 2.
+
+![](Rplot9.png)
+
+We calculated kappa statistics to estimate the inter-rater reliability for each of the STROBE items, these are presented in Table 1.
+
+```{r table1}
+
+doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
+
+if(doc.type == "docx"){pander(kappa_item, caption = "Table 1. Inter-rater reliability for the STROBE items")} else {kable(kappa_item)}
+
+```
+
+## Other reporting quality
+
+We collected reporting quality data in addition to the STROBE items.
+We intended to conduct exploratory analyses using these variables, for example comparing articles that reported using some form of reporting guidelines to those who did not.
+We did not conduct an exploratory analyses into the relationships between these variabels because subgroup analyses for all of these variables would have very small subgroup samples. 
+Instead we report the descriptive statistics for each variable in Table 2 below.
+
+*Table 2. Descriptive statistics for additional reporting quality variables*
+
+```{r table2, echo=F, include = F}
+
+###################
+# frequencies non-strobe ####
+##################
+
+# save version of not strobe df without qualtrics metadata variables to create table 2
+
+create_table2 <- function(){
+  # save pattern to find qualtrics variables
+  qual <- c("distribution_channel", "duration_in_seconds", "end_date", "finished", "progress","recorded_date", "response_id", "response_type", "start_date", "user_language")
+  qual_pat <- paste0("^", qual, ".", collapse = "|")
+
+  # find qualtrics columns in df
+  qual_cols <- grep(qual_pat, colnames(not_s_df))
+
+  # check there are 4 of each qual column
+  if(length(qual_cols) != 4* length(qual)) stop("some qualtrics cols missing")
+
+  table2 <- not_s_df[, -qual_cols]
+
+  # drop cols that are qualitative so contain unique info for each observation so no point in calculating frequencies
+  table2 <- table2[, -which(colnames(table2) %in% c("article_id", "correction_address", "open_access_address", "reg_id", "correction_address"))]
+
+  #convert all non-strobe cols into characters as easier than setting multiple different factor levels
+  for(i in colnames(table2)){
+    table2[[i]] <- as.character(table2[[i]])
+  }
+
+  # ensure UK and USA country values are capitalised as frequency is case-sensitive
+  uk_usa <- grep("^uk$|^usa$", table2$country, ignore.case = T)
+  table2$country[uk_usa] <- toupper(table2$country[uk_usa])
+
+  # drop cols that are characteristics
+  chars <- c("designs", "country", "year")
+  if(!all(chars %in% colnames(table2))) stop("missing char elements in colnames")
+  table2 <- table2[, -which(colnames(table2) %in% chars)]
+
+  # set missing & na as won't make a difference to absolute numbers
+  table2[is.na(table2)|table2 == "na"] <- "Not Applicable"
+  table2[table2 == "ni"] <- "No information"
+  table2[table2 == "TRUE"] <- "Yes"
+  table2[table2 == "FALSE"] <- "No"
+
+  # rename table2 colnames so will come out as full words in createtableon
+  colnames(table2)[colnames(table2) == "code"] <- "Report shared analysis code"
+  colnames(table2)[colnames(table2) == "exposure"] <- "Exposures reported as exposures"
+  colnames(table2)[colnames(table2) == "outcome"] <- "Outcomes reported as outcomes"
+  colnames(table2)[colnames(table2) == "reg"] <- "Report analyses registered"
+  colnames(table2)[colnames(table2) == "strobe"] <- "Report using STROBE guidelines"
+  colnames(table2)[colnames(table2) == "strobe_cite"] <- "Cite STROBE guidelines if used"
+  colnames(table2)[colnames(table2) == "ukb_credit"] <- "Report UK Biobank credit statement correctly"
+  colnames(table2)[colnames(table2) == "var_id"] <- "Report UK Biobank variable IDs"
+  colnames(table2)[colnames(table2) == "other_guidelines"] <- "Report using reporting guidelines other than STROBE"
+  colnames(table2)[colnames(table2) == "open_access_publisher"] <- "Published as open access"
+  colnames(table2)[colnames(table2) == "correction"] <- "Has a correction"
+  colnames(table2)[colnames(table2) == "ukb_app_pres"] <- "Report UK Biobank application number" 
+  colnames(table2)[colnames(table2) == "email_pres"] <- "Has email address for corresponding author"
+  colnames(table2)[colnames(table2) == "keywords_pres"] <- "Has keywords"
+  colnames(table2)[colnames(table2) == "coi_pres"] <- "Report conflict of interest statement"
+  colnames(table2)[colnames(table2) == "access_supp"] <- "Has supplementary material"
+
+# drop retraction col because none should be
+  table2 <- table2[, -which(colnames(table2) %in% "retraction")]
+  return(table2)
+}
+
+table2 <- create_table2()
+
+CreateTableOne(data = table2, includeNA = F) %>%
+  print(., noSpaces = T) %>%
+  write.csv(., "../outputs/table2.csv")
+
+```
 # Discussion
@@
 
+# Supplementary Material
+
+## Supplementary Table 2
+
+```{r SuppTable2}
+
+create_stab2 <- function(){
+  stab2 <- df[, which(colnames(df) %in% c(strobe_cols_names))]
+  # convert all to characters so can easily reset values and table will just treat as nominal
+  for(i in colnames(stab2)){
+    stab2[[i]] <- as.character(stab2[[i]])
+  }
+
+  # set missing to na as won't make a difference to absolute numbers
+  stab2[is.na(stab2)|stab2 == "na"] <- "Not Applicable"
+  stab2[stab2 == "PartiallyExternal"] <- "Partially-External"
+
+  # pivot so can table frequencies
+  stab2 <- tidyr::pivot_longer(stab2, colnames(stab2), names_to = "strobe_item", values_to = "value") %>%
+    # reorder so follows strobe order
+    .[mixedorder(as.character(.$strobe_item)), ]
+
+  # remove s from strobe item
+  stab2$strobe_item <- gsub("^s", "", stab2$strobe_item)
+
+  # table
+  stab2 <- table(stab2$strobe_item, stab2$value, useNA = "ifany")
+
+  return(stab2)
+}
+
+stab2 <- create_stab2()
+
+if(doc.type == "docx"){pander(stab2, caption = "Supplementary Table 2. Descriptive statistics for reporting quality of subdivisions of all STROBE items")} else {kable(stab2)}
+```
+
 # References

### 2021-11-05 manuscript/manuscript.Rmd
add some discussion

@@
   for(i in strobe_stem){
+    # convert stem into pattern so don't get greedy matches e.g. s2 match s22
+    pat <- paste0(c(paste0(i, "_"), paste0(i, "$")), collapse = "|")
     # save dataframe of variables for same strobe item so can sum rows
-    x <- df[,grep(i, colnames(df))]
+    x <- df[, grep(pat, colnames(df))]
     # create sum col name
@@
 
-Objectives
-
-Data sources
-
-Eligibility criteria
+The STrengthening the Reporting of OBservational studies in Epidemiology (STROBE)
+statement provides a reporting guideline for observational studies, including secondary analyses.
+We aimed to use the STROBE guidelines to assess the reporting quality of studies using UK Biobank data.
 
@@
 
-Intervention
-
-Study appraisal
-
-Synthesis methods
-
-Restuls
-
-Limitations
+70 cohort, cross-sectional and case-control studies published before 2019 were assessed using the STROBE guidelines.
+
+Results
+??? studies were identified, ??? were included and ??? were assessed.
 
 Conclusions
-
-Implications
 
@@
 ![](Rplot9.png)
+
+```{r assistance}
+
+bad <- as.data.frame(bar_data) %>%
+  .[order(as.character(.[,2]), .[,3]), ] %>%
+  .[-which(.$response == "Yes" & .$count > 30), ] %>%
+  .[-which(.$response == "No" & .$count < 20), ] %>%
+  .[-which(.$response == "Not Applicable"), ]
+
+good <- as.data.frame(bar_data) %>%
+  .[order(as.character(.[,2]), .[,3]), ] %>%
+  .[-which(.$response == "Yes" & .$count < 30), ] %>%
+  .[-which(.$response == "No" & .$count > 20), ] %>%
+  .[-which(.$response == "Not Applicable"), ]
+
+bad <- c("5", "8", "11", "12b", "12c", "16a", "19", "22", "12d", "13c", "14b", "16c", "1a", "9")
+
+good <- unique(bar_data$strobe_item[!bar_data$strobe_item %in% bad])
+```
+
+Figure 1 shows the best reported items are 2, 3, 4, 12a, and 18 with nearly all articles reporting each item fully.
+The good reporting of the main statistical methods suggested by 12a is encouraging for the reproducibility of the analyses.
+Most articles also fully reported item 21.
+
+Items 5, 8, 11, 12c, 16a, 19 and 22  were dominated by partial reporting.
+For some this is not a major concern.
+Items 5 and 8 relates to the setting, recruitment, location, dates, and assessment methods of the UK Biobank data collection.
+Although it does not meet the STROBE guidelines, all information for item 5 and 8 is fully reported in the publically available UK Biobank documentation.
+
+Other results indicate more uncertainty for the reproducibility of the included studies. 
+Authors poorly reported how they handled their variables (item 11) and missing data (item 12c).
+Another potential area of concern is the reporting of sensitivity, subgroup, interaction, and other analyses.
+Most reporting of the methods (items 12b and 12e) and results (item 17) from these analyses was incomplete. 
+Both of these issues could limit the reproducibility of UK Biobank analyses.
+
+Other items that were poorly reported do not pose the same concerns for reprodu
+In Figure 1, item 16a appears to be a potential cause for concern.
+Supplementary Table 2 shows that while nearly all authors fully report their confounder-adjusted estimates most are failing report their unadjusted elements and to justify why they included the confounders they did.
+This poses less threats to estimating reproducibility of the results but does suggest a widespread norm of researchers including confounders without a stated justification.
+Similarly the poor reporting of item 19 and 22 provide few threats to the reproducibility of the methods but suggest researchers are not expected to provide detailed information on potential bias or report the role of funders in their studies.
+
+The worst reported items were 12d, 13c, 14b and 16c, with most authors reporting no information about them.
+Again the implications are of varying concern for different items.
+Figure 1 shows few authors reported addressing loss to follow up in item 12d.
+This could be because many studies used NHS data linked to the UK Biobank as follow up data so the studies are likely to have low loss to follow up.
+Although confirming this will require further investigation into the individual studies.
+The impact of rare reporting of flow diagrams (item 13c) is also tempered by the good results for 13a and 13b, as it suggests authors do report the information just not in a flow diagram.
+Authors are poor at reporting absolute risk estimates (16c) and missing data (14c).
+
+```{r to-do}
+
+# CHECK OTHER STEMS ARE NOT BEING GREEDY
+# REMOVE ACCOUNT FOR SAMPLING FROM 12D LABEL?
+```
+
+**  12b - fv serious element so worrying
+
+lots of nos, lots of yes, few partial
+* 1a, 9,
 

### 2021-11-22 manuscript/manuscript.Rmd
manuscript draft 0.2 after RT comments, major changes to analysis too

@@
 library(lpSolve)
-library(kableExtra)
 library(pander)
@@
     # select those that are strobe items 
-    .[grep("s[[:digit:]]{1,2}", .$variable), ] %>%
+    .[grep(strobe_pattern, .$variable), ] %>%
     .[-grep("\\_ev|\\_star", .$variable), ]
@@
   ylab("Count") +
-  ggtitle("Figure 1. STROBE completion")+
   scale_fill_brewer(palette = "Greys") +
@@
 
+  # find other empty rows and add them to the na_stems
+  for(i in stem){
+    x <- df[, grep(i, colnames(df))]
+    i_na <- sum(is.na(x[[1]] != x[[2]]))
+    if(i_na == nrow(df)){
+      na_stem <- c(na_stem, i)
+    }
+  }
+
+  # keep only unique so not creating duplicates when add back in
+  na_stem <- unique(na_stem)
+
   # remove all stems that are duplicate or not applicable as will throw error in kappa2 if set "NA" to missing and all are set to NA after qualtrics
 
-  stem <- stem[-which(stem == na_stem)]
+  stem <- stem[-which(stem %in% na_stem)]
   #save empty objects so can save outputs from for loop
@@
   z <- c()
-  p_value <- c()
   variable <- c()
   n_match <- c()
-  n_conflicts <-c()
+  agree <-c()
   n_na <- c()
@@
     z <- c(z, y$statistic)
-    p_value <- c(p_value, y$p.value)
-    n_match <- c(n_match, sum(x[[1]] == x[[2]], na.rm = T))
-    n_conflicts <- c(n_conflicts, sum(x[[1]] != x[[2]], na.rm = T))
+    i_match <- sum(x[[1]] == x[[2]], na.rm = T)
+    n_match <- c(n_match, i_match)
     # count number of items with NA--NA dyads or notNA--NA dyads so are excluded from kappa
-    n_na <- c(n_na, sum(is.na(x[[1]] != x[[2]])))
+    i_na <- sum(is.na(x[[1]] != x[[2]]))
+    n_na <- c(n_na, i_na)
+    app <- sum(!is.na(x[[1]] != x[[2]]))
+    i_agree <- round((i_match / app) * 100, 0)
+    agree <- c(agree, i_agree)
   }
@@
   # bind objects into df so can render in manuscript
-  irr_df <- data.frame(STROBEItem = variable, Kappa = kappa, Zstatistic = z, pvalue = p_value, Matches = n_match, Conflicts = n_conflicts, NotApplicableUnsure = n_na)
+  irr_df <- data.frame(STROBEItem = variable, Kappa = kappa, Zstatistic = z, Matches = n_match, Agreement = agree, NotApplicableUnsure = n_na)
 
@@
 
+  irr_df$STROBEItem <- as.factor(irr_df$STROBEItem)
   # set rownames to null to ensure none created
@@
 }
+
+##########################
+# kappa for applicability #####
+##########################
 
@@
 #######################
+# calculate strobe scores for inter at subdivision level  ####
+########################
+
+inter_sub <- inter
+inter_sub[inter_sub == "NA" | inter_sub == "Unsure"] <- NA
+inter_sub[inter_sub == "Partially-External"] <- "Yes"
+
+# only select 
+inter_sub <- inter_sub[, grep(strobe_pattern, colnames(inter_sub))]
+kappa_sub <- create_kappa_df(inter_sub)
+
+#######################
 # calculate strobe scores for ternary for inter  ####
@@
   df[df == "NA" | df == "Unsure"] <- NA
-  strobe_cols_names <- grep("^s[[:digit:]]{1,2}", colnames(df), value =T)
+  strobe_cols_names <- grep(strobe_pattern, colnames(df), value =T)
 
@@
   for(i in strobe_stem){
-
+    # convert stem into pattern so don't get greedy matches e.g. s2 match s22
+    pat <- paste0(c(paste0(i, "_"), paste0(i, "$"), paste0(i, "\\.")), collapse = "|")
     # save dataframe of variables for same strobe item so can sum rows
-    x <- df[,grep(i, colnames(df))]
+    x <- df[,grep(pat, colnames(df))]
     # create sum col name
@@
     # sum rows in x if it is a dataframe 
-    if(!is.character(x)){
+    if(is.data.frame(x)){
       for(j in 1:nrow(x)){
-        if(all(x[j, ] %in% c("Yes", "Partially-External"), na.rm = T)) {
+        row <- as.character(x[j, ])[!is.na(as.character(x[j, ]))]
+        if(length(row) == 0) {
+          sum_var[j] <- NA
+        } else {
+          if(all(row %in% c("Yes", "Partially-External"), na.rm = T)) {
             sum_var[j] <- "Yes"
           } else {
-          if(!all(x[j, ] %in% c("Yes", "Partially-External"), na.rm = T) & !all(x[j, ] == "No", na.rm = T)) {
+            if(!all(row %in% c("Yes", "Partially-External"), na.rm = T) & !all(row == "No", na.rm = T)) {
               sum_var[j] <- "Partially"
             } else {
-            if(all(x[j, ] == "No", na.rm = T)) {
+              if(all(row == "No", na.rm = T)) {
                 sum_var[j] <- "No"
@@
                 stop(j, " in ", i, " doesn't contain just yes, pe, partially or no")
+              }
             }
@@
 
-kappa_item$STROBEItem <- gsub("^s|_.*|starred", "", kappa_item$STROBEItem)
-
 ```
@@
 statement provides a reporting guideline for observational studies, including secondary analyses.
-We aimed to use the STROBE guidelines to assess the reporting quality of studies using UK Biobank data.
+We used the STROBE guidelines to assess the reporting quality of studies using UK Biobank data.
 
@@
 
-70 cohort, cross-sectional and case-control studies published before 2019 were assessed using the STROBE guidelines.
+We assessed 70 cohort, cross-sectional and case-control studies published before 2019 using the STROBE guidelines.
 
 Results
-??? studies were identified, ??? were included and ??? were assessed.
+We identified ??? studies, included ???, and assessed a random sample of ???.
 
@@
 This accessibility offers researchers new opportunities but there is also uncertainty about how the nature of open data may impact the reliability of research findings. 
-One method by which the reliability of results from open data may be investigated is through replication. 
-Direct replication of secondary data analyses is aided by the publication of an original analysis plan, in which the authors describe their statistical analysis in full. 
-However, recent evidence suggests that it is unclear whether this is sufficient to ensure direct replicability of findings [@hardwicke2018; @naudet2018].
+One method researchers can use to investigate the reliability of results from open data is replication.
+In the case of secondary data analyses replications will involve new researchers retrieving the same dataset as the previous researchers and repeating their analyses.
+Researchers can make such direct replications of their work easier by publishing a detailed description of their statistical analysis.
+However, recent evidence suggests that it is unclear whether a detailed plan alone is sufficient to ensure direct replicability of findings [@hardwicke2018; @naudet2018].
 
@@
 The UKB is a prospective cohort study which collected a wide range of health data from 500,000 people aged 40-69 in the UK during 2006-2010 [@ukbiobankcoordinatingcentre2007].
-Importantly, UKB data is partially open data because it is only available to bona fide researchers conducting research in the public interest on a cost-recovery basis.
+Importantly, UKB data is "controlled data", meaning it is only available to bona fide researchers conducting research in the public interest on a cost-recovery basis.
 The UKB reserves the right to reject access applications for projects it deems inappropriate [@ukbiobankcoordinatingcentre2011].
-We first conducted a systematic review to (1) assess the completeness of UKB study reporting, (2) identify articles that are suitable for replication, and (3) inform reporting guidelines for other articles using the UKB and similar resources.
-We will evaluate reproducibility of studies conducted in the UKB given the abundance of published articles using data from this cohort and their contemporary nature, increasing the likelihood that authors will be contactable.
-Furthermore, they all use the same, professionally collected, maintained and diverse open resource. 
-This guarantees data accessibility and usability and means that differences between the original and replication results cannot be explained by poor data curation or heterogeneity between studies.
-
-Many previous systematic reviews of reporting standards across a range of disciplines used guidelines to assess the completeness of reporting [e.g. @agha2014; @cook2011; @plint2006).
+We plan to replicate UKB studies specifically because there is an abundance of published articles using the data and they are all contemporary, which increases the likelihood that authors will be contactable.
+Furthermore, they all use the same, professionally collected, maintained and diverse open resource which we will be able to access and use.
+
+We first conducted a systematic review to (1) assess the completeness of UKB study reporting, (2) identify articles that are suitable for replication, and (3) inform the development of reporting guidelines for other articles using the UKB and similar resources.
+Many previous systematic reviews of reporting quality across a range of disciplines used guidelines to assess the completeness of reporting [e.g. @agha2014; @cook2011; @plint2006).
 We used the STROBE (STrengthening the Reporting of Observational Studies in Epidemiology) Statement, which provides reporting guidelines for observational studies, to evaluate articles arising from UKB and to make comparisons between the articles.
 The STROBE Statement has been endorsed by over 100 journals and, as no major changes have been made to the original version published in 2007 [@cevallos2014], these guidelines are relevant to all studies conducted using the UKB.
-Given that studies conducted using UKB may be cohort, cross-sectional or ‘nested’ case-control, the review used the combined STROBE Statement for cohort, cross-sectional and case-control designs, and its extensions, to assess the reporting standards in UKB articles.
-
-We aimed to conduct a systematic review of articles reporting studies using data from the UKB with the aim of answering the following questions:
+Given that studies conducted using UKB may be cohort, cross-sectional or ‘nested’ case-control, the review used the combined STROBE Statement for cohort, cross-sectional and case-control designs, and its extensions, to assess the reporting quality in UKB articles.
+
+We conducted a systematic review of articles reporting studies using data from the UKB with the aim of answering the following questions:
 
@@
 
-Eligible articles are full research articles that report analyses of UKB data. Each article that is excluded at the full-text level of screening will be listed with the reason for exclusion.
+Eligible articles were full research articles that reported analyses of UKB data. 
+We give the reason for exclusion for each article we excluded at the full-text level of screening.
 
@@
 * Publication date: After 30 March 2012. 
-This is the date the UKB was launched and UKB data was first available beforehand.
+This is the date the UKB was launched and UKB data was first available for researchers to use.
 * Publication type: Full research articles (including simulation and modelling articles).
 * Language: English. 
-Relevant studies written in another language may be eligible but, to ensure accurate and precise evaluation, the article would require a professional translation which is not feasible given the review’s resources.
+Relevant studies written in another language may be relevant but, to ensure accurate and precise evaluation, the article would require a professional translation which is not feasible given the review’s resources.
 * Data source: Article’s findings were produced from analysis(es) of data from the UKB. 
@@
 * Retracted articles
+* Studies with inaccessible supplementary material. 
+We was to ensure our estimates of reporting completeness included the entire version of record.
 
@@
 
-Four databases were electronically searched on 15/01/2019 for UKB articles: PubMed, EMBASE, Web of Science Core Collection (WoSCC) and PsycINFO. 
+We searched four databases on 15/01/2019 for UKB articles: PubMed, EMBASE, Web of Science Core Collection (WoSCC) and PsycINFO. 
 The search strategy was limited to articles published from 2012 onwards and there was no restriction on language or publication type.
@@
 
-## study selection
+## Study selection
 
 We conducted four stages of screening for eligibility.
-First, using Endnote, KD removed duplicate results, checked the results for completeness, updated the metadata if necessary, and retrieved the full-texts of all results.
+First, using Endnote, KD removed duplicate results, checked the results for completeness, updated the metadata if necessary, and retrieved the full-texts of all articles.
+KD updated the metadata by checking each article's Endnote metadata contained the article's DOI, and adding them if they were missing.
 KD checked completeness using a list of published full research articles that have used the UKB Resource, provided by the UKB on 27 November 2018. 
+We conducted a systematic search despite having this list because we wished to capture all articles from 2018.
 Using the full-texts of the results, KD assessed if results were full research articles, and excluded reviews, corrections, conference abstracts, etc.
 
-Second, the results were imported into Rayyan [@ouzzani2016] and screened by BW and KD to determine whether they met the inclusion criteria.
-If insufficient information was available to determine a paper’s eligibility, the authors were contacted to request the required information. 
-If this information was unobtainable, the article was excluded.
+Second, we imported the results into Rayyan [@ouzzani2016]. 
+BW and KD screened them to determine if they met the inclusion criteria.
+If insufficient information was available to determine a paper’s eligibility, we contacted the authors and requested the required information, if it was unobtainable we excluded the article.
 Time constraints meant we excluded some articles with insufficient information without contacting the authors.
@@
 Third, We classified studies as "traditional epidemiology", "Mendelian randomisation" and "Other", according to the type of data they analysed.
-After being piloted on 20 studies, the classification criteria were:
+After we piloted these classifications on 20 studies, the classification criteria were:
 
@@
 
-Articles abstracts were classified independently by MG and KD. 
-Full texts were examined where classification is unclear from the abstract.
-Any conflicts were discussed by KD and BW and resolved by mutual consent if possible. 
-If not resolved after discussion RR assign the category.
+BW and KD independently classified articles' abstracts.
+They examined full texts if classification was unclear from the abstract alone.
+They discussed any conflicts and resolved them by mutual consent if possible, if not resolved after discussion then RR assigned the category.
 Articles could be allocated into more than one classification group.
@@
 * Studies incorrectly classified as observational epidemiological studies.
-* Studies with inaccessible supplementary material
-* Studies we did consider to be observiational epidemiological studies but could not consider them to have cohort, case-control or cross-sectional designs - for example those using prediction models
+* Studies with inaccessible supplementary material. 
+* Studies we did consider to be observational epidemiological studies but could not consider them to have cohort, case-control or cross-sectional designs - for example those using prediction models
 
@@
 To ensure coders completed the same STROBE items for each article KD and MG independently identified the study designs used by each of the 178 “observational epidemiology” articles by completing the Study Design Form for each article.
-In the form, extractors also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
-Any conflicts were discussed by the two extractors and resolved by mutual consent if possible.
+In the form, coders also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
+Any conflicts were discussed by the two coders and resolved by mutual consent if possible.
 If not resolved after discussion, RR determined the study designs. 
@@
 The articles were randomly assigned an ID number from 1-178.
-The articles with an ID number between 1 and 80 were selected for the first round of data extraction. 
-KD was assigned as the first extractor to all 80 articles. 
-The second extractor was randomly assigned from MG, RR or BW.
+We selected the articles with an ID number between 1 and 80 for the first round of data extraction. 
+We assigned KD as the first coder for all 80 articles and randomly assigned MG, RR, or BW as the second coder.
 We initially randomly selected 80 articles with the hope that we would randomly select more studies once data extraction for the first 80 were complete.
-Unfortunately we did not have time to assess more studies.
+Unfortunately we did not have time to assess more studies so we do not have data on the remaining 98 articles. 
+Figure 1 shows the flow of records identified from our searches and reasons for exclusion.
+
+![Figure 1. Flow diagram of included articles](flow.png)
 
@@
 
-The Data Extraction Form was hosted on Qualtrics and piloted twice. 
+The data extraction form was hosted on Qualtrics. 
 This form combined the data items described as part of the "Data Extraction Form" and "Reporting Quality Assessment" in the protocol, but not all were kept.
-In the first, 5 observational epidemiology articles were piloted by MG, KD and/or RR. 
-Major changes were made so the Data Extraction Form will be completed again for the articles in the first pilot. 
+We conducted two pilots.
+In the first, MG, KD and/or RR piloted 5 observational epidemiology articles.
+We made major changes to the Data Extraction Form so we completed it again for the articles in the first pilot. 
 In the second, KD piloted the changed form on 3 articles. 
-No changes were made so the responses for articles in the second pilot were retained.
-
-The extractors independently completed the Data Extraction Form for each article assigned to them.
-Initially, any conflicts were discussed by the two extractors and resolved by mutual consent if possible. 
+We made no major changes so the we retained the responses for articles in the second pilot.
+See the PDF version of the data extraction form for full information on our variables and how we measured them at https://osf.io/hngcj/.
+
+The coders independently completed the data extraction form for each article assigned to them.
+Almost all variables were categorical so coders assigned the most appropriate category for each variable.
+Variables derived from the STROBE items had the same six categories:
+
+* "NA = this information was not relevant to any of the studies in the article"
+* "Unsure = I am unsure what to respond    "
+* "No = this information was relevant but none of it was present"
+* "Partially = some of the information was present"
+* "Partially-External = some of the information was present and/or the authors report that additional information can be found in an external source (i.e.
+not in the article or supplementary material)"
+* "Yes = all of the information was present".
+
+After two coders completed one data extraction form for each article, the two coders discussed any conflicts and resolved them by mutual consent if possible. 
 If not resolved after discussion, RR resolved the conflict.
-However, given the number of conflicts this could not be sustained.
-To reduce the time-burden on other authors KD and RR devised a series of "rules" and general guidelines KD could use to resolve conflicts by herself. 
-This means the majority of conflicts were resolved by these "Rules" or KD's own judgement.
-An explanation of the rules and guidelines can be found at https://osf.io/hngcj/.
+However, given the number of conflicts, this practice could not be sustained.
+To reduce the time-burden on other authors KD and RR devised a series of "rules" and general guidelines KD could use to resolve conflicts by themself. 
+This means KD resolved the majority of conflicts by these "Rules".
+For example, for item 1a of the STROBE checklist we agreed that authors describing a study as a "longitudinal" is not sufficient according to the STROBE guidelines but it still gives some indication of the study design.
+Therefore, we agreed on the rule "Rule = "Partially" if "longitudinal" referred to" for item 1a.
+The rules and guidelines can be found at https://osf.io/hngcj/.
 
@@
 
-The items in Table 1 were extracted manually by KD or automatically by Endnote which contained the article’s metadata exported from the database the article was located in.
+The items in Table 2 were extracted manually by KD or automatically by Endnote which contained the article’s metadata exported from the database the article was located in.
 
@@
 If the missing information is not contained in a linked publication we planned to contact the article’s authors.
-However, we realised this would give an inaccurate assessment of the verision of record. 
-Also few articles had linked publications and that there was a lot of missing or unclear information.
+However, we realised this would give an inaccurate assessment of the version of record. 
+Also few articles had linked publications and there was a lot of missing or unclear information.
 Therefore, we did not attempt to retrieve any missing items in the Data Extraction Form, from either the authors or linked articles.
 
+## Data analysis
+
+Using the data extraction form on Qualtrics, we collected data on the reporting quality on numerous variables, including those derived from the STROBE items.
+This mean our raw data estimated the reporting quality for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it. 
+For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor, this was primarily because raters could not agree when the items were or were not applicable.
+In the data extraction form raters could indicate items as "NA", "Unsure", "No", "Partially", "Partially-External" or "Yes".
+After resolving conflicts, no "Unsure" ratings remained. 
+We collapsed the "Partially-External" into the "Yes" catergory.
+This was because we agreed reporting there were very few "Partially-External" categories after resolving conflicts, to make the kappa statistics more interpretable, and we understood that reporting where the required information can be found could be considered complete reporting.
+
+Some STROBE items were written as multiple-barrelled questions e.g " Report assessment methods for outcomes, exposures, predictors, confounder, and effect modifiers". 
+We subdivided these questions into separate variables in the data extraction form.
+We could have assessed each item without dividing it up but collecting the reporting completeness at the subdivision level we have a more granular understanding of what information is or is not reported.
+We summed the subdivisions back together to calculate the reporting completeness of each STROBE item they corresponded to. 
+This gave us an understanding of the reporting quality for the entire STROBE item and reduced the number of variables from `r length(unique(stab2$strobe_item))` to `r length(unique(bar_data$strobe_item))`.
+
+We calculated Cohen's kappa statistics to estimate the inter-rater reliability for each of the STROBE items.
+We treated any "Not Applicable" or "Unsure" values as missing data in the kappa calculations.  
+For our subgroup analyses we repeated the analyses for the subdivisions.
+We conducted all our data cleaning and analysis in R.
+
 # Results
 
-## Study selection
-
 ## Study characteristics
 
-Of the `r nrow(df)` included studies `r sum(grepl("Case-control", df$designs))` contained case-control designs, `r sum(grepl("Cohort", df$designs))` cohort, and `r sum(grepl("Cross-sectional", df$designs))` cross-sectional.
+Of the `r nrow(df)` included articles `r sum(grepl("Case-control", df$designs))` contained case-control designs, `r sum(grepl("Cohort", df$designs))` cohort, and `r sum(grepl("Cross-sectional", df$designs))` cross-sectional.
 These numbers add up to more than 70 because some articles contained cross-sectional and cohort studies.
 UK based researchers led the vast majority with `r sum(grepl("UK", df$country, ignore.case = F))` having authors affiliated with institutions in the UK.
+The mean number of authors was `r round(mean(as.numeric(as.character(df$authors))))` (SD = `r round(sd(as.numeric(as.character(df$authors))))`).
 `r sum(grepl("2014", df$year))` studies were published in 2014, `r sum(grepl("2015", df$year))` in 2015, `r sum(grepl("2016", df$year))` in 2016, `r sum(grepl("2017", df$year))` in 2017, `r sum(grepl("2018", df$year))` in 2018, and `r sum(grepl("2019", df$year))` in 2019.
@@
 
-Using the data extraction form on Qualtrics, we assessed reporting quality separately for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it.
-For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor and the applicability of the items was often uncertain. 
-We then summed the subdivisions to indicate the reporting quality of each STROBE item, these results are presented in Figure 1. 
-For completeness, the results for the subdivisions are available in Supplementary Table 2.
-
-We originally assessed reporting quality separately for all subdivisions of the STROBE items.
-The results for the subdivisions are reported in Supplementary Table 2.
-
-![](Rplot9.png)
-
-```{r assistance}
-
-bad <- as.data.frame(bar_data) %>%
-  .[order(as.character(.[,2]), .[,3]), ] %>%
-  .[-which(.$response == "Yes" & .$count > 30), ] %>%
-  .[-which(.$response == "No" & .$count < 20), ] %>%
-  .[-which(.$response == "Not Applicable"), ]
-
-good <- as.data.frame(bar_data) %>%
-  .[order(as.character(.[,2]), .[,3]), ] %>%
-  .[-which(.$response == "Yes" & .$count < 30), ] %>%
-  .[-which(.$response == "No" & .$count > 20), ] %>%
-  .[-which(.$response == "Not Applicable"), ]
-
-bad <- c("5", "8", "11", "12b", "12c", "16a", "19", "22", "12d", "13c", "14b", "16c", "1a", "9")
-
-good <- unique(bar_data$strobe_item[!bar_data$strobe_item %in% bad])
+Figure 1 presents the reporting completeness for each STROBE item, to ensure a comprehensive understanding of our results, Supplementary Table 2 provides the reporting completeness for each subdivision.
+
+![Figure 2. STROBE completion](Rplot9.png)
+
+```{r to-do}
+
+# CHECK OTHER STEMS ARE NOT BEING GREEDY
+# REMOVE ACCOUNT FOR SAMPLING FROM 12D LABEL in figure 1?
 ```
 
-Figure 1 shows the best reported items are 2, 3, 4, 12a, and 18 with nearly all articles reporting each item fully.
-The good reporting of the main statistical methods suggested by 12a is encouraging for the reproducibility of the analyses.
-Most articles also fully reported item 21.
-
-Items 5, 8, 11, 12c, 16a, 19 and 22  were dominated by partial reporting.
-For some this is not a major concern.
-Items 5 and 8 relates to the setting, recruitment, location, dates, and assessment methods of the UK Biobank data collection.
-Although it does not meet the STROBE guidelines, all information for item 5 and 8 is fully reported in the publically available UK Biobank documentation.
-
-Other results indicate more uncertainty for the reproducibility of the included studies. 
-Authors poorly reported how they handled their variables (item 11) and missing data (item 12c).
-Another potential area of concern is the reporting of sensitivity, subgroup, interaction, and other analyses.
-Most reporting of the methods (items 12b and 12e) and results (item 17) from these analyses was incomplete. 
-Both of these issues could limit the reproducibility of UK Biobank analyses.
-
-Other items that were poorly reported do not pose the same concerns for reprodu
-In Figure 1, item 16a appears to be a potential cause for concern.
-Supplementary Table 2 shows that while nearly all authors fully report their confounder-adjusted estimates most are failing report their unadjusted elements and to justify why they included the confounders they did.
-This poses less threats to estimating reproducibility of the results but does suggest a widespread norm of researchers including confounders without a stated justification.
-Similarly the poor reporting of item 19 and 22 provide few threats to the reproducibility of the methods but suggest researchers are not expected to provide detailed information on potential bias or report the role of funders in their studies.
-
-The worst reported items were 12d, 13c, 14b and 16c, with most authors reporting no information about them.
-Again the implications are of varying concern for different items.
-Figure 1 shows few authors reported addressing loss to follow up in item 12d.
-This could be because many studies used NHS data linked to the UK Biobank as follow up data so the studies are likely to have low loss to follow up.
-Although confirming this will require further investigation into the individual studies.
-The impact of rare reporting of flow diagrams (item 13c) is also tempered by the good results for 13a and 13b, as it suggests authors do report the information just not in a flow diagram.
-Authors are poor at reporting absolute risk estimates (16c) and missing data (14c).
-
-```{r to-do}
-
-# CHECK OTHER STEMS ARE NOT BEING GREEDY
-# REMOVE ACCOUNT FOR SAMPLING FROM 12D LABEL?
+Table 1 presents Cohen's kappa statistics for each of the STROBE items.
+Many items had poor inter-rater reliability with `r sum(kappa_item$Kappa < 0.2 & kappa_item$Kappa != 0, na.rm = T)` items having a kappa below 0.2.
+The main reason for this is that the data extraction form was subjective.
+Conflicts occured when one coder did not notice that the required information was reported and when coders interpretted the STROBE item differently.
+The list of "rules" we compiled give an indication for conflicts that occurred more than once, how we decided to resolve them and our justification for it.
+
+```{r table1}
+
+pander(kappa_item, caption = "Table 1. Inter-rater reliability for the STROBE items")
+
 ```
 
-**  12b - fv serious element so worrying
-
-lots of nos, lots of yes, few partial
-* 1a, 9,
-
-We calculated kappa statistics to estimate the inter-rater reliability for each of the STROBE items, these are presented in Table 1.
-
-```{r table1}
-
-doc.type <- knitr::opts_knit$get('rmarkdown.pandoc.to')
-
-if(doc.type == "docx"){pander(kappa_item, caption = "Table 1. Inter-rater reliability for the STROBE items")} else {kable(kappa_item)}
-
-```
-
 ## Other reporting quality
@@
 We collected reporting quality data in addition to the STROBE items.
-We intended to conduct exploratory analyses using these variables, for example comparing articles that reported using some form of reporting guidelines to those who did not.
-We did not conduct an exploratory analyses into the relationships between these variabels because subgroup analyses for all of these variables would have very small subgroup samples. 
+We intended to conduct exploratory analyses using these variables to answer our second research question, for example comparing articles that reported using some form of reporting guidelines to those who did not.
+We did not conduct an exploratory analyses into the relationships between these variables because subgroup analyses for all of these variables would have very small subgroup samples. 
 Instead we report the descriptive statistics for each variable in Table 2 below.
@@
 
+  # replace unclear value to make it more understandable
+  table2[table2 == "No available elsewhere"] <- "No but publicly available on another platform"
+  table2[table2 == "yes_exact"] <- "Yes - exact wording"
+  table2[table2 == "yes_almost"] <- "Yes - almost exact wording"
+
   # rename table2 colnames so will come out as full words in createtableon
   colnames(table2)[colnames(table2) == "code"] <- "Report shared analysis code"
-  colnames(table2)[colnames(table2) == "exposure"] <- "Exposures reported as exposures"
-  colnames(table2)[colnames(table2) == "outcome"] <- "Outcomes reported as outcomes"
+  colnames(table2)[colnames(table2) == "exposure"] <- "Exposure variables clearly labelled as exposures"
+  colnames(table2)[colnames(table2) == "outcome"] <- "Outcome variables clearly labeled as outcomes"
   colnames(table2)[colnames(table2) == "reg"] <- "Report analyses registered"
@@
 
+Our results indicate that the STROBE items best reported by UK Biobank studies are those regarding backround (item 2), objectives (item 3), explaining design early (item 4), main statistical methods (item 12a), and summarising results (item 18) with nearly all articles reporting each item fully.
+The good reporting of the main statistical methods is encouraging for the potential reproducibility of the analysis methods.
+Most articles also fully reported information on generalisability (item 21).
+
+Partial reporting dominated items regarding setting, recruitment, location, and dates (item 5); assessment methods (item 8); variable analysis and grouping (item 11); addressing missing data (item 12c); main results and definitions of confounders (item 16a); limitations and bias (item 19); and funder details (item 22).
+For some items partial reporting is not a major concern.
+Although it does not meet the STROBE guidelines, all information for item 5 and 8 is fully reported in the publicly available UK Biobank documentation.
+For others, partial reporting casts doubt on the ability of new researchers to know how to repeat the included studies. 
+Authors rarely fully reported how they handled their variables (item 11) and missing data (item 12c).
+Another potential area of concern is the reporting of sensitivity, subgroup, interaction, and other analyses.
+Most reporting of the methods (items 12b and 12e) and results (item 17) from these analyses was incomplete. 
+Both of these issues could limit how accurately the UK Biobank analysis methods can be repeated.
+
+Other items that were poorly reported carry different concerns to those of reproducibility.
+Most authors partially reported item 16a. 
+Supplementary Table 2 shows that while nearly all authors fully report their confounder-adjusted estimates most are failing report their unadjusted elements and to justify why they included the confounders they did.
+This information may be less vital to reproducing analyses than other items, but it does suggest a widespread norm of researchers including confounders without a stated justification.
+Similarly the significant lack of reporting for item 9, 19 and 22 suggests it is acceptable for researchers not to provide detailed analysis of potential bias in their studies or report the role of funders in their studies.
+Lastly, many authors failing to mention the design of their study in their title or abstract (item 1a) could create problems in retrieving studies during searches.
+
+The worst reported items were 12d, 13c, 14b and 16c, with most authors reporting no information about them.
+Again the implications are of varying concern for different items.
+Figure 2 shows few authors reported addressing loss to follow up in item 12d.
+This could be because many studies used NHS data linked to the UK Biobank as follow up data so the studies are likely to have low loss to follow up.
+Although confirming this will require further investigation into the individual studies.
+The impact of rare reporting of flow diagrams (item 13c) is also tempered by the good results for 13a and 13b, as it suggests authors do report the information just not in a flow diagram.
+Authors are poor at reporting absolute risk estimates (16c) and missing data (14c).
+
+We coded additional variables to measure reporting quality of information not covered by the STROBE guidelines.
+62 articles reported a conflict of interest statement, 41 provided their studies' UK Biobank application number, and 59 reported the UK Biobank credit statement correctly.
+However, the reporting was poor for the variables we included to assess the reproducibility of the analysis methods.
+Only 2 papers shared any part of their analysis code, 2 reported having registered their analyses though whether these registrations were before or after seeing the data was unclear, and 10 reported at least one UK Biobank variable ID.
+
 ## Limitations
 
+* appropriateness of strobe for secondary data analysis - e.g. item 10 excluded completely
+* kappa statistics -- limitations of double coding for something so subjective
+* deviations from systematic review method & protocol
+* reverse engineering of reporting guidelines = inappropriate to use as reporting assessment?
+  ** Releasing ‘adherence checklists’ with reporting guidelines would help 
+reporting assessments
+
 ## Conclusions
+
+Good reporting of statistical methods will aid replications of the main results. 
+Comparing replication results to the original may be difficult given the poor reporting of unadjusted estimates.
+The lack of shared code will also make replications more difficult.
 
@@
 
+## Supplementary Table 1
+
+```{r STable1, echo= F}
+
+create_stab1 <- function(){
+  # read in rules to add into table
+  stab1 <- read.csv("../data/rules.csv", encoding = "UTF-8", na.strings = c(""), stringsAsFactors = F)
+
+  # save ids of all variable cols
+  cols_id <- grep("variable", colnames(stab1))
+
+  # save variables element so can save for loop products
+  x <- c()
+
+  # paste all variable cols together and save products
+  for(i in 1:nrow(stab1)){
+    x <- c(x, paste(stab1[i, cols_id], collapse = ","))
+  }
+
+  # clean variable names to strobe items and remove all missing items and spaces
+  x <- gsub(",NA.*| |starred|^s", "", x) %>%
+    gsub(",s", ",", .) %>%
+    # add spaces back in so more readable
+    gsub(",", ", ", .)
+
+  # add in as column
+  stab1$STROBEItem <- x
+
+  colnames(stab1)[colnames(stab1) == "rule"] <- "Rule"
+  colnames(stab1)[colnames(stab1) == "STROBEItem"] <- "STROBE Item" 
+  # remove all unnecessary cols
+  stab1 <- stab1[, -grep("variable|response", colnames(stab1))]
+
+  stab1$Rule <- gsub("^Rule = |^Rule =", "", stab1$Rule)
+
+  return(stab1)
+
+}
+
+stab1 <- create_stab1()
+
+pander(stab1, caption = "Supplementary Table 1. Rules for resolving conflicts", justify = c("left", "center"))
+```
+
 ## Supplementary Table 2
 
-```{r SuppTable2}
+```{r STable2, echo=F}
+
+# create labels for strobe items at subdivision level
+
+sub_labels <- read.csv("../outputs/extraction_dict.csv", stringsAsFactors = F, encoding = "UTF-8") %>%
+  .[grep(strobe_pattern, .$variable), ] %>%
+  .[-grep("_ev$", .$variable),]
+
+# create stabe
 
@@
   stab2[is.na(stab2)|stab2 == "na"] <- "Not Applicable"
-  stab2[stab2 == "PartiallyExternal"] <- "Partially-External"
+  stab2[stab2 == "PartiallyExternal"] <- "Yes"
 
@@
   # remove s from strobe item
-  stab2$strobe_item <- gsub("^s", "", stab2$strobe_item)
-
-  # table
-  stab2 <- table(stab2$strobe_item, stab2$value, useNA = "ifany")
+  stab2$strobe_item <- gsub("^s|starred", "", stab2$strobe_item)
+
+  strobe_levels <-  unique(stab2$strobe_item)
+
+  # set levels as mixed order so order preserved in ggplot and pander
+  stab2$strobe_item<- factor(stab2$strobe_item, levels =  strobe_levels[gtools::mixedorder(strobe_levels)])
+
+  # order levels 
+  stab2$value <- as.factor(stab2$value)
+  levels(stab2$value) <- list(NotApplicable = "Not Applicable", No = "No", Partially = "Partially", Yes = "Yes")
 
@@
 stab2 <- create_stab2()
-
-if(doc.type == "docx"){pander(stab2, caption = "Supplementary Table 2. Descriptive statistics for reporting quality of subdivisions of all STROBE items")} else {kable(stab2)}
+stab2_tab <- table(stab2$strobe_item, stab2$value, useNA = "ifany")
+
+png("Rplot10.png", width = 1300, height = 1500)
+
+# strobe item on x axis and plot % yes
+stab2 %>%ggplot(aes(x=strobe_item, y=..count.., fill = value)) + 
+  # blue fill
+  geom_bar() +
+  theme_classic() +
+  theme(text = element_text(size=30))+
+  #scale_x_discrete(labels= labels$x_labels) +
+  #scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
+  coord_flip() +
+  xlab("STROBE Item") +
+  ylab("Count") +
+  scale_fill_brewer(palette = "Greys") +
+  #geom_text(aes(label = count), vjust = 0, size = 8) +
+  theme(plot.title = element_text(size = 22)) 
+dev.off()
+
+pander(stab2_tab, caption = "Supplementary Table 2. Descriptive statistics for reporting quality of subdivisions of all STROBE items")
 ```
 
+## Supplementary Table 3
+
+```{r, echo = F}
+pander(kappa_sub, caption = "Supplementary Table 3. Kappa statistics for reporting quality of subdivisions of all STROBE items")
+```
+
 # References

### 2021-11-22 rscripts/02_data_extraction/03_merge_resolved_conflicts.R
add number of authors, save res_pat as enviro object so can reuse accurately, set arbitrary limit on unique values

@@ -369,16 +369,18 @@ df <- merge_all_three()
 # clean strobe responses ####
 ####################
 
-# catergories should only be a few check this
+# save pattern of naming format used to indicate a resolved conflict for a strobe item so can clean them>
+# these will be responses to strobe items with the name of the coder who resolved them after the value >
+# only becky, I, or both of us resolved them so it will be "becky" or "me"
+res_pat <- paste0(c(" becky ",  "becky$", " me ", " me$", " us ", "us$"), collapse = "|") 
 
 find_all_values <- function(df){
-  # remove all columns that aren't strobe variables (i.e. don't begin with s[digit])
-  df <- df[, grep("^s[0-9]", colnames(df))]
   ls <- list()
   for(i in colnames(df)){
     ls[[i]] <- unique(df[[i]])
   }
-  ls <- ls[lengths(ls) != 33]
+  # don't find values for cols with more than 20 unique values (arbitrary boundary) as these will be qualitative variables
+  ls <- ls[lengths(ls) < 20]
   ls <- lapply(ls, `length<-`, max(lengths(ls)))
   x <- as.data.frame(ls)
   all_value <- c()
@@ -389,7 +391,7 @@ find_all_values <- function(df){
 }
 
 clean_values <- function(){
-  
+  # find all values 
   vals <- find_all_values(df)
   rules<- sort(unique(grep("rule", vals, value = T, ignore.case = T))) %>%
     # remove any text that comes before rule, this will clean any like 'NA - resolved but Rule ...'
@@ -413,19 +415,15 @@ clean_values <- function(){
     df[, i] <- col
   }
   
-  
-  # clean non-strobe items
+  # clean all strobe values that are resolved conflicts but don't have a rule
   
   for(i in colnames(df)){
-    # save values that indicate a resolved conflict for a strobe item so can clean them>
-    # these will be responses to strobe items with the name of the coder who resolved them after the value >
-    # only becky, I, or both of us resolved them so it will be "becky" or "me"
-    res <- c(" becky ",  "becky$", " me ", " me$", " us ", "us$") 
-    # save as a pattern so can use in gsub
-    pat <- paste(res, collapse = "|")
     x <- df[[i]]
-    if(any(grepl(pat, x, ignore.case = T))){
-      ids <- grep(pat, x, ignore.case = T)
[…]

### 2021-11-29 manuscript/manuscript.Rmd
fix flow diagram, move code for supplementary material up so all code is in same chuck at start so all objects for in-text r code will exist, small corrections, move deviations from protocol supplementary material, drop supp table 2 & change supp table 1 to figure & update references accordingly, add limitations

@@
   Studies'
-author: Emma Anderson, George Davey Smith, Katie Drax, Mark Gibson, Marcus Munafò,
-  Benjamin Woolf, and Rebecca Richmond
+author: Emma Anderson, George Davey Smith, Katie Drax, Mark Gibson, Marcus Munafò, Robert Thibault, Benjamin Woolf, and Rebecca Richmond
 date: "`r format(Sys.time(), '%d %B %Y')`"
@@
 
-```
-
-# Abstract
-
-Background
-
-The STrengthening the Reporting of OBservational studies in Epidemiology (STROBE)
-statement provides a reporting guideline for observational studies, including secondary analyses.
-We used the STROBE guidelines to assess the reporting quality of studies using UK Biobank data.
-
-Participants
-
-We assessed 70 cohort, cross-sectional and case-control studies published before 2019 using the STROBE guidelines.
-
-Results
-We identified ??? studies, included ???, and assessed a random sample of ???.
-
-Conclusions
-
-```{r, prisma}
-
-# 4. questions picos
-
-```
-
-# Background
-
-There is a general move towards making data freely accessible, or ‘open’, by academic and governmental institutions alike [@vasilevsky2017; @worldwidewebfoundation2018]. 
-Completely open data allows any individual, regardless of their motivation, discipline or skillset, to access and analyse a dataset for free. 
-This accessibility offers researchers new opportunities but there is also uncertainty about how the nature of open data may impact the reliability of research findings. 
-One method researchers can use to investigate the reliability of results from open data is replication.
-In the case of secondary data analyses replications will involve new researchers retrieving the same dataset as the previous researchers and repeating their analyses.
-Researchers can make such direct replications of their work easier by publishing a detailed description of their statistical analysis.
-However, recent evidence suggests that it is unclear whether a detailed plan alone is sufficient to ensure direct replicability of findings [@hardwicke2018; @naudet2018].
-
-To investigate the reliability of findings from open data, we will attempt to replicate all eligible research articles that use UK Biobank (UKB) data. 
-The UKB is a prospective cohort study which collected a wide range of health data from 500,000 people aged 40-69 in the UK during 2006-2010 [@ukbiobankcoordinatingcentre2007].
-Importantly, UKB data is "controlled data", meaning it is only available to bona fide researchers conducting research in the public interest on a cost-recovery basis.
-The UKB reserves the right to reject access applications for projects it deems inappropriate [@ukbiobankcoordinatingcentre2011].
-We plan to replicate UKB studies specifically because there is an abundance of published articles using the data and they are all contemporary, which increases the likelihood that authors will be contactable.
-Furthermore, they all use the same, professionally collected, maintained and diverse open resource which we will be able to access and use.
-
-We first conducted a systematic review to (1) assess the completeness of UKB study reporting, (2) identify articles that are suitable for replication, and (3) inform the development of reporting guidelines for other articles using the UKB and similar resources.
-Many previous systematic reviews of reporting quality across a range of disciplines used guidelines to assess the completeness of reporting [e.g. @agha2014; @cook2011; @plint2006).
-We used the STROBE (STrengthening the Reporting of Observational Studies in Epidemiology) Statement, which provides reporting guidelines for observational studies, to evaluate articles arising from UKB and to make comparisons between the articles.
-The STROBE Statement has been endorsed by over 100 journals and, as no major changes have been made to the original version published in 2007 [@cevallos2014], these guidelines are relevant to all studies conducted using the UKB.
-Given that studies conducted using UKB may be cohort, cross-sectional or ‘nested’ case-control, the review used the combined STROBE Statement for cohort, cross-sectional and case-control designs, and its extensions, to assess the reporting quality in UKB articles.
-
-We conducted a systematic review of articles reporting studies using data from the UKB with the aim of answering the following questions:
-
-1. What is the quality of reporting of findings from the UKB (in terms of the detail, clarity and completeness with which authors report their design, analysis and results)?
-2. Does reporting quality vary across the type of studies conducted?
-
-# Methods
-
-## Protocol registration
-
-This study was preregistered before we began to screen the search results [@drax2019].
-
-## Eligibility criteria
-
-Eligible articles were full research articles that reported analyses of UKB data. 
-We give the reason for exclusion for each article we excluded at the full-text level of screening.
-
-Inclusion criteria:
-
-* Publication date: After 30 March 2012. 
-This is the date the UKB was launched and UKB data was first available for researchers to use.
-* Publication type: Full research articles (including simulation and modelling articles).
-* Language: English. 
-Relevant studies written in another language may be relevant but, to ensure accurate and precise evaluation, the article would require a professional translation which is not feasible given the review’s resources.
-* Data source: Article’s findings were produced from analysis(es) of data from the UKB. 
-This includes studies which reported findings from other studies alongside findings from UKB data.
-
-Exclusion criteria:
-
-* Article findings were produced from analysis(es) of pooled data from multiple studies.
-These are studies which have used data from participants from the UKB and other datasets in the same analysis.
-* Meta-analyses; narrative and systematic reviews; preprints; post-prints; replies; letters to the editor; corrections; and any other publication types that are not full research articles.
-* Retracted articles
-* Studies with inaccessible supplementary material. 
-We was to ensure our estimates of reporting completeness included the entire version of record.
-
-## Information sources and search strategy
-
-We searched four databases on 15/01/2019 for UKB articles: PubMed, EMBASE, Web of Science Core Collection (WoSCC) and PsycINFO. 
-The search strategy was limited to articles published from 2012 onwards and there was no restriction on language or publication type.
-
-The full search strategy for each database is included in Appendix A of the protocol [osf ref]. 
-To summarise, the terms "UK Biobank", "UKB", "UKBiobank", and "UKB Resource" were included in the "Title", "Abstract", "Keywords", and "All Fields" fields or equivalent of each database.
-
-```{r}
-# is there counts for the number of articles from th eUKB list?
-
-```
-
-## Study selection
-
-We conducted four stages of screening for eligibility.
-First, using Endnote, KD removed duplicate results, checked the results for completeness, updated the metadata if necessary, and retrieved the full-texts of all articles.
-KD updated the metadata by checking each article's Endnote metadata contained the article's DOI, and adding them if they were missing.
-KD checked completeness using a list of published full research articles that have used the UKB Resource, provided by the UKB on 27 November 2018. 
-We conducted a systematic search despite having this list because we wished to capture all articles from 2018.
-Using the full-texts of the results, KD assessed if results were full research articles, and excluded reviews, corrections, conference abstracts, etc.
-
-Second, we imported the results into Rayyan [@ouzzani2016]. 
-BW and KD screened them to determine if they met the inclusion criteria.
-If insufficient information was available to determine a paper’s eligibility, we contacted the authors and requested the required information, if it was unobtainable we excluded the article.
-Time constraints meant we excluded some articles with insufficient information without contacting the authors.
-
-Third, We classified studies as "traditional epidemiology", "Mendelian randomisation" and "Other", according to the type of data they analysed.
-After we piloted these classifications on 20 studies, the classification criteria were:
-
-*	Traditional epidemiology – studies investigating associations between exposures and health outcomes using the UK Biobank data alone. 
-*	Mendelian randomisation – studies that conduct Mendelian randomisation analyses on UK Biobank data alone
-*	Other – articles that do not meet the ‘Traditional epidemiology’ or ‘MR’ classification criteria described above or include analyses of genetic or imaging data.
-
-BW and KD independently classified articles' abstracts.
-They examined full texts if classification was unclear from the abstract alone.
-They discussed any conflicts and resolved them by mutual consent if possible, if not resolved after discussion then RR assigned the category.
-Articles could be allocated into more than one classification group.
-We then excluded all but the traditional epidemiology studies.
-The Mendialian randomisation studies were not relevant to this paper but we included it as a classification because the data may be used for one of our author's future projects.
-
-Only including observational epidemiological studies is a major departure from the protocol in which we stated that studies would be "classified according to their data type" during data extraction but that all classifications would be included.
-This was so we could analyse articles according to their appropriate STROBE extension, such as the strengthening the reporting of genetic association studies (STREGA) statement [@little2009].
-Given that it would be too time-consuming to extract data from all studies in all included articles and to design multiple data extraction forms, we decided to extract data and assess the reporting quality of studies that could be assessed using the original STROBE guidelines only, i.e.
-traditional epidemiology studies.
-
-Fourth, we catergorised the 178 studies classified as observational epidemiological studies into those with cohort, case-control and cross-sectional study designs using the Study Design Form described below.
-During this process KD also examined the notices attached to the online version of the observational epidemiology articles to identify any retracted articles. 
-None were identified.
-However, we did identify studies that should be excluded, these studies were:
-
-* Studies incorrectly classified as observational epidemiological studies.
-* Studies with inaccessible supplementary material. 
-* Studies we did consider to be observational epidemiological studies but could not consider them to have cohort, case-control or cross-sectional designs - for example those using prediction models
-
-## Data collection process
-
-### Study Design Form
-
-Different STROBE items are relevant for cohort, case-control and cross-sectional study designs. 
-To ensure coders completed the same STROBE items for each article KD and MG independently identified the study designs used by each of the 178 “observational epidemiology” articles by completing the Study Design Form for each article.
-In the form, coders also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
-Any conflicts were discussed by the two coders and resolved by mutual consent if possible.
-If not resolved after discussion, RR determined the study designs. 
-The Study Design Form was hosted on Qualtrics but a PDF version can be found at https://osf.io/jfk24/.
-
-The articles were randomly assigned an ID number from 1-178.
-We selected the articles with an ID number between 1 and 80 for the first round of data extraction. 
-We assigned KD as the first coder for all 80 articles and randomly assigned MG, RR, or BW as the second coder.
-We initially randomly selected 80 articles with the hope that we would randomly select more studies once data extraction for the first 80 were complete.
-Unfortunately we did not have time to assess more studies so we do not have data on the remaining 98 articles. 
-Figure 1 shows the flow of records identified from our searches and reasons for exclusion.
-
-![Figure 1. Flow diagram of included articles](flow.png)
-
-### Data Extraction Form
-
-The data extraction form was hosted on Qualtrics. 
-This form combined the data items described as part of the "Data Extraction Form" and "Reporting Quality Assessment" in the protocol, but not all were kept.
-We conducted two pilots.
-In the first, MG, KD and/or RR piloted 5 observational epidemiology articles.
-We made major changes to the Data Extraction Form so we completed it again for the articles in the first pilot. 
-In the second, KD piloted the changed form on 3 articles. 
-We made no major changes so the we retained the responses for articles in the second pilot.
-See the PDF version of the data extraction form for full information on our variables and how we measured them at https://osf.io/hngcj/.
-
-The coders independently completed the data extraction form for each article assigned to them.
-Almost all variables were categorical so coders assigned the most appropriate category for each variable.
-Variables derived from the STROBE items had the same six categories:
-
-* "NA = this information was not relevant to any of the studies in the article"
-* "Unsure = I am unsure what to respond    "
-* "No = this information was relevant but none of it was present"
-* "Partially = some of the information was present"
-* "Partially-External = some of the information was present and/or the authors report that additional information can be found in an external source (i.e.
-not in the article or supplementary material)"
-* "Yes = all of the information was present".
-
-After two coders completed one data extraction form for each article, the two coders discussed any conflicts and resolved them by mutual consent if possible. 
-If not resolved after discussion, RR resolved the conflict.
-However, given the number of conflicts, this practice could not be sustained.
-To reduce the time-burden on other authors KD and RR devised a series of "rules" and general guidelines KD could use to resolve conflicts by themself. 
-This means KD resolved the majority of conflicts by these "Rules".
-For example, for item 1a of the STROBE checklist we agreed that authors describing a study as a "longitudinal" is not sufficient according to the STROBE guidelines but it still gives some indication of the study design.
-Therefore, we agreed on the rule "Rule = "Partially" if "longitudinal" referred to" for item 1a.
-The rules and guidelines can be found at https://osf.io/hngcj/.
-
-### Other data collection
-
-The items in Table 2 were extracted manually by KD or automatically by Endnote which contained the article’s metadata exported from the database the article was located in.
-
-### Unclear and Missing Information.
-
-In our protocol we planned to locate and examine publications linked to the article, e.g.
-related full research articles, comments, corrections, etc, so that we could retrieve any missing items in the Data Extraction Form from them.
-If the missing information is not contained in a linked publication we planned to contact the article’s authors.
-However, we realised this would give an inaccurate assessment of the version of record. 
-Also few articles had linked publications and there was a lot of missing or unclear information.
-Therefore, we did not attempt to retrieve any missing items in the Data Extraction Form, from either the authors or linked articles.
-
-## Data analysis
-
-Using the data extraction form on Qualtrics, we collected data on the reporting quality on numerous variables, including those derived from the STROBE items.
-This mean our raw data estimated the reporting quality for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it. 
-For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor, this was primarily because raters could not agree when the items were or were not applicable.
-In the data extraction form raters could indicate items as "NA", "Unsure", "No", "Partially", "Partially-External" or "Yes".
-After resolving conflicts, no "Unsure" ratings remained. 
-We collapsed the "Partially-External" into the "Yes" catergory.
-This was because we agreed reporting there were very few "Partially-External" categories after resolving conflicts, to make the kappa statistics more interpretable, and we understood that reporting where the required information can be found could be considered complete reporting.
-
-Some STROBE items were written as multiple-barrelled questions e.g " Report assessment methods for outcomes, exposures, predictors, confounder, and effect modifiers". 
-We subdivided these questions into separate variables in the data extraction form.
-We could have assessed each item without dividing it up but collecting the reporting completeness at the subdivision level we have a more granular understanding of what information is or is not reported.
-We summed the subdivisions back together to calculate the reporting completeness of each STROBE item they corresponded to. 
-This gave us an understanding of the reporting quality for the entire STROBE item and reduced the number of variables from `r length(unique(stab2$strobe_item))` to `r length(unique(bar_data$strobe_item))`.
-
-We calculated Cohen's kappa statistics to estimate the inter-rater reliability for each of the STROBE items.
-We treated any "Not Applicable" or "Unsure" values as missing data in the kappa calculations.  
-For our subgroup analyses we repeated the analyses for the subdivisions.
-We conducted all our data cleaning and analysis in R.
-
-# Results
-
-## Study characteristics
-
-Of the `r nrow(df)` included articles `r sum(grepl("Case-control", df$designs))` contained case-control designs, `r sum(grepl("Cohort", df$designs))` cohort, and `r sum(grepl("Cross-sectional", df$designs))` cross-sectional.
-These numbers add up to more than 70 because some articles contained cross-sectional and cohort studies.
-UK based researchers led the vast majority with `r sum(grepl("UK", df$country, ignore.case = F))` having authors affiliated with institutions in the UK.
-The mean number of authors was `r round(mean(as.numeric(as.character(df$authors))))` (SD = `r round(sd(as.numeric(as.character(df$authors))))`).
-`r sum(grepl("2014", df$year))` studies were published in 2014, `r sum(grepl("2015", df$year))` in 2015, `r sum(grepl("2016", df$year))` in 2016, `r sum(grepl("2017", df$year))` in 2017, `r sum(grepl("2018", df$year))` in 2018, and `r sum(grepl("2019", df$year))` in 2019.
-
-## STROBE Reporting quality
-
-Figure 1 presents the reporting completeness for each STROBE item, to ensure a comprehensive understanding of our results, Supplementary Table 2 provides the reporting completeness for each subdivision.
-
-![Figure 2. STROBE completion](Rplot9.png)
-
-```{r to-do}
-
-# CHECK OTHER STEMS ARE NOT BEING GREEDY
-# REMOVE ACCOUNT FOR SAMPLING FROM 12D LABEL in figure 1?
-```
-
-Table 1 presents Cohen's kappa statistics for each of the STROBE items.
-Many items had poor inter-rater reliability with `r sum(kappa_item$Kappa < 0.2 & kappa_item$Kappa != 0, na.rm = T)` items having a kappa below 0.2.
-The main reason for this is that the data extraction form was subjective.
-Conflicts occured when one coder did not notice that the required information was reported and when coders interpretted the STROBE item differently.
-The list of "rules" we compiled give an indication for conflicts that occurred more than once, how we decided to resolve them and our justification for it.
-
-```{r table1}
-
-pander(kappa_item, caption = "Table 1. Inter-rater reliability for the STROBE items")
-
-```
-
-## Other reporting quality
-
-We collected reporting quality data in addition to the STROBE items.
-We intended to conduct exploratory analyses using these variables to answer our second research question, for example comparing articles that reported using some form of reporting guidelines to those who did not.
-We did not conduct an exploratory analyses into the relationships between these variables because subgroup analyses for all of these variables would have very small subgroup samples. 
-Instead we report the descriptive statistics for each variable in Table 2 below.
-
-*Table 2. Descriptive statistics for additional reporting quality variables*
-
-```{r table2, echo=F, include = F}
+kappa_item$STROBEItem <-   gsub("_sum|starred|^s", "", kappa_item$STROBEItem)
 
 ###################
-# frequencies non-strobe ####
+# create table 2 - frequencies non-strobe ####
 ##################
@@
 
+#############
+# create stab2####
+############
+
+# create labels for strobe items at subdivision level
+
+sub_labels <- read.csv("../outputs/extraction_dict.csv", stringsAsFactors = F, encoding = "UTF-8") %>%
+  .[grep(strobe_pattern, .$variable), ] %>%
+  .[-grep("_ev$", .$variable),]
+
+# create stabe
+
+create_stab2 <- function(){
+  stab2 <- df[, which(colnames(df) %in% c(strobe_cols_names))]
+  # convert all to characters so can easily reset values and table will just treat as nominal
+  for(i in colnames(stab2)){
+    stab2[[i]] <- as.character(stab2[[i]])
+  }
+
+  # set missing to na as won't make a difference to absolute numbers
+  stab2[is.na(stab2)|stab2 == "na"] <- "Not Applicable"
+  stab2[stab2 == "PartiallyExternal"] <- "Yes"
+
+  # pivot so can table frequencies
+  stab2 <- tidyr::pivot_longer(stab2, colnames(stab2), names_to = "strobe_item", values_to = "value") %>%
+    # reorder so follows strobe order
+    .[mixedorder(as.character(.$strobe_item)), ]
+
+  strobe_levels <-  unique(stab2$strobe_item)
+
+  # set levels as mixed order so order preserved in ggplot and pander
+  stab2$strobe_item<- factor(stab2$strobe_item, levels =  strobe_levels[gtools::mixedorder(strobe_levels)])
+
+  # order levels 
+  stab2$value <- as.factor(stab2$value)
+  levels(stab2$value) <- list(NotApplicable = "Not Applicable", No = "No", Partially = "Partially", Yes = "Yes")
+
+  return(stab2)
+}
+
+labels <-  read.csv("../outputs/extraction_dict.csv", encoding = "UTF-8", na.strings = c(""), stringsAsFactors = T)
+
+stab2 <- create_stab2()
+stab2_tab <- table(stab2$strobe_item, stab2$value, useNA = "ifany")
+
+stab2 <- left_join(stab2, labels, by = c("strobe_item" = "variable"))
+
+stab2$question <- gsub("Indicate if the authors report the following items for all the studies in the article \\(unless a study design is specified\\) - |Indicate if the authors report the following items for all the cross-sectional studies in the article - |_ - |Indicate if the authors report the following items for all the case-control studies in the article - |Indicate if the authors report the following items for all the cohort studies in the article - |Indicate if the authors report the following items for all the cross-sectional and cohort studies in the article - ", "", stab2$question)
+
+png("Rplot10.png", width = 1300, height = 1500)
+
+# strobe item on x axis and plot % yes
+stab2 %>%ggplot(aes(x=strobe_item, y=..count.., fill = value)) + 
+  # blue fill
+  geom_bar() +
+  theme_classic() +
+  theme(text = element_text(size=30))+
+ # scale_x_discrete(labels= unique(stab2$question)) +
+  #scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
+  coord_flip() +
+  xlab("STROBE Item") +
+  ylab("Count") +
+  scale_fill_brewer(palette = "Greys") +
+  #geom_text(aes(label = count), vjust = 0, size = 8) +
+  theme(plot.title = element_text(size = 22)) 
+dev.off()
+
 ```
+
+``` {r}
+# Abstract
+#
+#Background
+#
+#The STrengthening the Reporting of OBservational studies in #Epidemiology (STROBE)
+#statement provides a reporting guideline for observational #studies, including secondary analyses.
+#We used the STROBE guidelines to assess the reporting quality #of studies using UK Biobank data.
+#
+#Participants
+#
+#We assessed 70 cohort, cross-sectional and case-control #studies published before 2019 using the STROBE guidelines.
+#
+#Results
+#We identified ??? studies, included ???, and assessed a #random sample of ???.
+#
+#Conclusions
+```
+
+```{r, prisma}
+
+# 4. questions picos
+
+```
+
+# Background
+
+There is a general move towards making data freely accessible, or ‘open’, by academic and governmental institutions alike [@vasilevsky2017; @worldwidewebfoundation2018]. 
+Completely open data allows any individual, regardless of their motivation, discipline or skillset, to access and analyse a dataset for free. 
+This accessibility offers researchers new opportunities but there is also uncertainty about how the nature of open data may impact the reliability of research findings. 
+One method researchers can use to investigate the reliability of results from open data is replication.
+In the case of secondary data analyses replications will involve new researchers retrieving the same dataset as the previous researchers and repeating their analyses.
+Researchers can make such direct replications of their work easier by publishing a detailed description of their statistical analysis.
+However, recent evidence suggests that it is unclear whether a detailed plan alone is sufficient to ensure direct replicability of findings [@hardwicke2018; @naudet2018].
+
+To investigate the reliability of findings from open data, we will attempt to replicate all eligible research articles that use UK Biobank (UKB) data. 
+The UKB is a prospective cohort study which collected a wide range of health data from 500,000 people aged 40-69 in the UK during 2006-2010 [@ukbiobankcoordinatingcentre2007].
+Importantly, UKB data is "controlled data", meaning it is only available to bona fide researchers conducting research in the public interest on a cost-recovery basis.
+The UKB reserves the right to reject access applications for projects it deems inappropriate [@ukbiobankcoordinatingcentre2011].
+We plan to replicate UKB studies specifically because there is an abundance of published articles using the data and they are all contemporary, which increases the likelihood that authors will be contactable.
+Furthermore, they all use the same, professionally collected, maintained and diverse open resource which we will be able to access and use.
+
+We first conducted a systematic review to (1) assess the completeness of UKB study reporting, (2) identify articles that are suitable for replication, and (3) inform the development of reporting guidelines for other articles using the UKB and similar resources.
+Many previous systematic reviews of reporting quality across a range of disciplines used guidelines to assess the completeness of reporting [e.g. @agha2014; @cook2011; @plint2006).
+We used the STROBE (STrengthening the Reporting of Observational Studies in Epidemiology) Statement, which provides reporting guidelines for observational studies, to evaluate articles arising from UKB and to make comparisons between the articles.
+The STROBE Statement has been endorsed by over 100 journals and, as no major changes have been made to the original version published in 2007 [@cevallos2014], these guidelines are relevant to all studies conducted using the UKB.
+Given that studies conducted using UKB may be cohort, cross-sectional or ‘nested’ case-control, the review used the combined STROBE Statement for cohort, cross-sectional and case-control designs, and its extensions, to assess the reporting quality in UKB articles.
+
+We conducted a systematic review of articles reporting studies using data from the UKB with the aim of answering the following questions:
+
+1. What is the quality of reporting of findings from the UKB (in terms of the detail, clarity and completeness with which authors report their design, analysis and results)?
+2. Does reporting quality vary across the type of studies conducted?
+
+# Methods
+
+## Protocol registration
+
+This study was preregistered before we began to screen the search results [@drax2019].
+For a description of the deviations from the registered protocol see the Supplementary Material.
+
+## Inclusion criteria
+
+The inclusion criteria were:
+
+* Publication date: After 30 March 2012. 
+This is the date the UKB was launched and UKB data was first available for researchers to use.
+* Publication type: Full research articles (including simulation and modelling articles).
+* Language: English. 
+Relevant studies written in another language may be relevant but, to ensure accurate and precise evaluation, the article would require a professional translation which is not feasible given the review’s resources.
+* Data source: Article’s findings were produced from analysis(es) of data from the UKB. 
+This includes studies which reported findings from other studies alongside findings from UKB data.
+* Study type: Observational epidemiology studies
+* Study design: Cross-sectional, cohort, or case-control study design.
+
+The exclusion criteria were:
+
+* Article findings were produced from analysis(es) of pooled data from multiple studies.
+These are studies which have used data from participants from the UKB and other datasets in the same analysis.
+* Meta-analyses; narrative and systematic reviews; preprints; post-prints; replies; letters to the editor; corrections; and any other publication types that are not full research articles.
+* Retracted articles
+* Studies with inaccessible supplementary material. 
+This was to ensure our estimates of reporting completeness included the entire version of record.
+
+## Information sources and search strategy
+
+We searched four databases on 15/01/2019 for UKB articles: PubMed, EMBASE, Web of Science Core Collection (WoSCC) and PsycINFO. 
+The search strategy was limited to articles published from 2012 onwards and there was no restriction on language or publication type.
+
+The full search strategy for each database is included in Appendix A of the protocol [@drax2019]. 
+To summarise, the terms "UK Biobank", "UKB", "UKBiobank", and "UKB Resource" were included in the "Title", "Abstract", "Keywords", and "All Fields" fields or equivalent of each database.
+
+```{r}
+# is there counts for the number of articles from th eUKB list?
+
+```
+
+## Study selection
+
+We conducted three stages of screening for eligibility.
+First, using Endnote, KD removed duplicate results, checked the results for completeness, updated the metadata if necessary, and retrieved the full-texts of all articles.
+KD updated the metadata by checking each article's Endnote metadata contained the article's DOI, and adding them if they were missing.
+KD checked completeness using a list of published full research articles that have used the UKB Resource, provided by the UKB on 27 November 2018. 
+We conducted a systematic search despite having this list because we wished to capture all articles from 2018.
+Using the full-texts of the results, KD assessed if results were full research articles, and excluded reviews, corrections, conference abstracts, etc.
+
+Second, we imported the results into Rayyan [@ouzzani2016]. 
+BW and KD independently screened the results and indicated if they met the inclusion criteria.
+At this point we included all articles whether or not they had a cross-sectional, cohort, or case-control study design because the "Study design" inclusion critera was added later.
+They examined full texts if inclusion was unclear from the abstract alone.
+They discussed any conflicts and resolved them by mutual consent if possible, if not resolved after discussion then RR assigned the category.
+If insufficient information was available in the full text to determine a paper’s eligibility, we contacted the authors and requested the required information, if it was unobtainable we excluded the article.
+Time constraints meant we excluded XXX articles with insufficient information without contacting the authors.
+
+## Data collection process
+
+### Study Design Form
+
+Different STROBE items are relevant for cohort, case-control and cross-sectional study designs. 
+To ensure coders completed the same STROBE items for each article KD and MG independently identified the study designs used by each of the 178 “observational epidemiology” articles by completing the Study Design Form for each article.
+In the form, coders also indicated if all supplementary material can be accessed via University of Bristol subscriptions. 
+Any conflicts were discussed by the two coders and resolved by mutual consent if possible.
+If not resolved after discussion, RR determined the study designs. 
+The Study Design Form was hosted on Qualtrics but a PDF version can be found at https://osf.io/jfk24/.
+During this process we also identified studies we did consider to be observational epidemiological studies but could not consider them to have cohort, case-control or cross-sectional designs - for example those using prediction models.
+Any articles we could not classify as cohort, case-control, or cross-sectional were excluded at this point.
+
+The articles were randomly assigned an ID number from 1-178.
+We selected the articles with an ID number between 1 and 80 for the first round of data extraction. 
+We assigned KD as the first coder for all 80 articles and randomly assigned MG, RR, or BW as the second coder.
+We initially randomly selected 80 articles with the hope that we would randomly select more studies once data extraction for the first 80 were complete.
+Unfortunately we did not have time to assess more studies so we do not have data on the remaining 98 articles. 
+Figure 1 shows the flow of records identified from our searches and reasons for exclusion.
+
+![Figure 1. Flow diagram of included articles](flow.png)
+
+### Data Extraction Form
+
+Two coders independently completed the data extraction form for each article assigned to them.
+Almost all variables in the form were categorical so coders assigned the most appropriate category for each variable.
+We conducted two pilots.
+In the first, MG, KD and/or RR piloted 5 observational epidemiology articles.
+We made major changes to the Data Extraction Form so we completed it again for the articles in the first pilot. 
+In the second, KD piloted the changed form on 3 articles. 
+We made no major changes so we kept the responses for articles in the second pilot.
+The PDF version of the data extraction form contains full information on our variables and how we measured them and is available at https://osf.io/hngcj/.
+
+The form contained questions to measure variables relating to the reporting quality of different information.
+We derived some variables from the STROBE guidelines.
+Some STROBE items were written as multiple-barrelled questions e.g " Report assessment methods for outcomes, exposures, predictors, confounder, and effect modifiers". 
+We subdivided these questions into separate variables in the data extraction form.
+We could have assessed each item without dividing it up but collecting the reporting completeness at the subdivision level we have a more granular understanding of what information is or is not reported.
+
+Variables derived from the STROBE items had the same six categories:
+
+* "NA = this information was not relevant to any of the studies in the article"
+* "Unsure = I am unsure what to respond    "
+* "No = this information was relevant but none of it was present"
+* "Partially = some of the information was present"
+* "Partially-External = some of the information was present and/or the authors report that additional information can be found in an external source (i.e.
+not in the article or supplementary material)"
+* "Yes = all of the information was present".
+
+After two coders completed one data extraction form for each article, the two coders discussed any conflicts and resolved them by mutual consent if possible. 
+If not resolved after discussion, RR resolved the conflict.
+However, given the number of conflicts, this practice could not be sustained.
+To reduce the time-burden on other authors KD and RR devised a series of "rules" and general guidelines KD could use to resolve conflicts by themself. 
+This means KD resolved the majority of conflicts by these "Rules".
+For example, for item 1a of the STROBE checklist we agreed that authors describing a study as a "longitudinal" is not sufficient according to the STROBE guidelines but it still gives some indication of the study design.
+Therefore, we agreed on the rule "Rule = "Partially" if "longitudinal" referred to" for item 1a.
+The rules and guidelines can be found at https://osf.io/hngcj/.
+
+### Other data collection
+
+The items in Table 2 were extracted manually by KD or automatically by Endnote which contained the article’s metadata exported from the database the article was located in.
+
+### Unclear and Missing Information.
+
+In our protocol we planned to locate and examine publications linked to the article, e.g.
+related full research articles, comments, corrections, etc, so that we could retrieve any missing items in the Data Extraction Form from them.
+If the missing information is not contained in a linked publication we planned to contact the article’s authors.
+However, we realised this would give an inaccurate assessment of the version of record. 
+Also few articles had linked publications and there was a lot of missing or unclear information.
+Therefore, we did not attempt to retrieve any missing items in the Data Extraction Form, from either the authors or linked articles.
+
+## Data analysis
+
+Using the data extraction form on Qualtrics, we collected data on the reporting quality on numerous variables, including those derived from the STROBE items.
+This mean our raw data estimated the reporting quality for all subdivisions of the STROBE items and for the “starred” items associated with any STROBE items that had an asterisk (*) next to it. 
+For the final analysis, we excluded all starred subdivisions because inter-rater reliability was so poor, this was primarily because raters could not agree when the items were or were not applicable.
+In the data extraction form raters could indicate items as "NA", "Unsure", "No", "Partially", "Partially-External" or "Yes".
+After resolving conflicts, no "Unsure" ratings remained. 
+We collapsed the "Partially-External" into the "Yes" catergory.
+This was because we agreed reporting there were very few "Partially-External" categories after resolving conflicts, to make the kappa statistics more interpretable, and we understood that reporting where the required information can be found could be considered complete reporting.
+We summed the subdivisions back together to calculate the reporting completeness of each STROBE item they corresponded to. 
+This gave us an understanding of the reporting quality for the entire STROBE item and reduced the number of variables from `r length(unique(stab2$strobe_item))` to `r length(unique(bar_data$strobe_item))`.
+
+We calculated Cohen's kappa statistics to estimate the inter-rater reliability for each of the STROBE items.
+We treated any "Not Applicable" or "Unsure" values as missing data in the kappa calculations.  
+For our subgroup analyses we repeated the analyses for the subdivisions.
+We conducted all our data cleaning and analysis in R.
+
+# Results
+
+## Study characteristics
+
+Of the `r nrow(df)` included articles `r sum(grepl("Case-control", df$designs))` contained case-control designs, `r sum(grepl("Cohort", df$designs))` cohort, and `r sum(grepl("Cross-sectional", df$designs))` cross-sectional.
+These numbers add up to more than 70 because some articles contained cross-sectional and cohort studies.
+UK based researchers led the vast majority with `r sum(grepl("UK", df$country, ignore.case = F))` having authors affiliated with institutions in the UK.
+The mean number of authors was `r round(mean(as.numeric(as.character(df$authors))))` (SD = `r round(sd(as.numeric(as.character(df$authors))))`).
+`r sum(grepl("2014", df$year))` studies were published in 2014, `r sum(grepl("2015", df$year))` in 2015, `r sum(grepl("2016", df$year))` in 2016, `r sum(grepl("2017", df$year))` in 2017, `r sum(grepl("2018", df$year))` in 2018, and `r sum(grepl("2019", df$year))` in 2019.
+
+## STROBE Reporting quality
+
+Figure 1 presents the reporting completeness for each STROBE item, to ensure a comprehensive understanding of our results, Supplementary Figure 1 provides the reporting completeness for each subdivision.
+
+![Figure 2. STROBE completion](Rplot9.png)
+
+```{r to-do}
+
+# CHECK OTHER STEMS ARE NOT BEING GREEDY
+# REMOVE ACCOUNT FOR SAMPLING FROM 12D LABEL in figure 1?
+```
+
+Table 1 presents Cohen's kappa statistics for each of the STROBE items.
+Many items had poor inter-rater reliability with `r sum(kappa_item$Kappa < 0.2 & kappa_item$Kappa != 0, na.rm = T)` items having a kappa below 0.2.
+The main reason for this is that the data extraction form was subjective.
+Conflicts occured when one coder did not notice that the required information was reported and when coders interpretted the STROBE item differently.
+The list of "rules" we compiled give an indication for conflicts that occurred more than once, how we decided to resolve them and our justification for it.
+
+```{r table1}
+
+pander(kappa_item, caption = "Table 1. Inter-rater reliability for the STROBE items")
+
+```
+
+## Other reporting quality
+
+We collected reporting quality data in addition to the STROBE items.
+We intended to conduct exploratory analyses using these variables to answer our second research question, for example comparing articles that reported using some form of reporting guidelines to those who did not.
+We did not conduct an exploratory analyses into the relationships between these variables because subgroup analyses for all of these variables would have very small subgroup samples. 
+Instead we report the descriptive statistics for each variable in Table 2 below.
+
+*Table 2. Descriptive statistics for additional reporting quality variables*
+
 # Discussion
@@
 Most authors partially reported item 16a. 
-Supplementary Table 2 shows that while nearly all authors fully report their confounder-adjusted estimates most are failing report their unadjusted elements and to justify why they included the confounders they did.
+Supplementary Figure 1 shows that while nearly all authors fully report their confounder-adjusted estimates most are failing report their unadjusted elements and to justify why they included the confounders they did.
 This information may be less vital to reproducing analyses than other items, but it does suggest a widespread norm of researchers including confounders without a stated justification.
@@
 
-* appropriateness of strobe for secondary data analysis - e.g. item 10 excluded completely
-* kappa statistics -- limitations of double coding for something so subjective
-* deviations from systematic review method & protocol
-* reverse engineering of reporting guidelines = inappropriate to use as reporting assessment?
-  ** Releasing ‘adherence checklists’ with reporting guidelines would help 
-reporting assessments
+Our study has several limitations.
+Firstly, many variables in the data extraction form we used to assess reporting quality had poor inter-rater reliability.
+The kappa statistics, use of "Rules" to solve conflicts, and exclusion of the starred subdivisions suggest as much.
+This questions the appropriateness of our data extraction form, and the STROBE guidelines it is a based on, as a measure of reporting quality since it does not seem to accurately estimate reporting quality across coders.
+This poor reliability may be because the STROBE guidelines were not designed as a measure of reporting quality.
+The STROBE guidelines are reporting guidelines designed for authors to use to check their manuscript contains all the relevant information, while they are writing up.
+The guidelines are not designed as an assessment of reporting by people other than the authors.
+Reverse engineering the guidelines for a purpose they were not designed for may be why the inter-rater reliability is poor.
+The appropriateness of the STROBE guidelines for secondary data analyses, such as those using the UK Biobank, is also questionable.
+We used the STROBE guidelines to create our data extraction form because it was the most appropriate reporting guidelines available.
+However, the STROBE guidelines did not necessarily capture the information we needed, leading to us excluding several STROBE items and including additional variables.
+For example, the guidelines were written in 2007, therefore there is no question on whether or how authors reported analysis code.
+We also excluded several items or subdivisions of items because they requested reporting of the same information or information not applicable to a seondary data analysis, specifically item 10, two subdivisions of item 6a, one subdivision of item 5, and one subdivision of item 8.
+See the rule guidelines for further information on these duplicate and inapplicable variables (https://osf.io/hngcj/).
 
@@
 
+```{r}
+#Third, using Rayyan again, we classified studies as "traditional epidemiology", "Mendelian randomisation" and "Other", according to the #type of data they analysed.
+#After we piloted these classifications on 20 studies, the classification criteria were:
+#
+#*	Traditional epidemiology – studies investigating associations between exposures and health outcomes using the UK Biobank data alone. 
+#*	Mendelian randomisation – studies that conduct Mendelian randomisation analyses on UK Biobank data alone
+#*	Other – articles that do not meet the ‘Traditional epidemiology’ or ‘MR’ classification criteria described above or include analyses of #genetic or imaging data.
+#
+#BW and KD independently classified articles' abstracts.
+#They examined full texts if classification was unclear from the abstract alone.
+#They discussed any conflicts and resolved them by mutual consent if possible, if not resolved after discussion then RR assigned the #category.
+#Articles could be allocated into more than one classification group.
+#We then excluded all but the traditional epidemiology studies.
+#The Mendialian randomisation studies were not relevant to this paper but we included it as a classification because the data may be used for one of our author's future projects.
+
+```
+
+Only including observational epidemiological studies is a major departure from the protocol in which we stated that studies would be "classified according to their data type" during data extraction but that all classifications would be included.
+This was so we could analyse articles according to their appropriate STROBE extension, such as the strengthening the reporting of genetic association studies (STREGA) statement [@little2009].
+Given that it would be too time-consuming to extract data from all studies in all included articles and to design multiple data extraction forms, we decided to extract data and assess the reporting quality of studies that could be assessed using the original STROBE guidelines only, i.e.
+traditional epidemiology studies.
+
+The final data extraction form combined the data items described as part of the "Data Extraction Form" and "Reporting Quality Assessment" in the protocol, but not all were kept.
+
+## Supplementary Figure 1
+
+*Supplementary Figure 1. Descriptive statistics for reporting quality of subdivisions of all STROBE items*
+![](Rplot10.png)
+
 ## Supplementary Table 1
 
-```{r STable1, echo= F}
-
-create_stab1 <- function(){
-  # read in rules to add into table
-  stab1 <- read.csv("../data/rules.csv", encoding = "UTF-8", na.strings = c(""), stringsAsFactors = F)
-
-  # save ids of all variable cols
-  cols_id <- grep("variable", colnames(stab1))
-
-  # save variables element so can save for loop products
-  x <- c()
-
-  # paste all variable cols together and save products
-  for(i in 1:nrow(stab1)){
-    x <- c(x, paste(stab1[i, cols_id], collapse = ","))
-  }
-
-  # clean variable names to strobe items and remove all missing items and spaces
-  x <- gsub(",NA.*| |starred|^s", "", x) %>%
-    gsub(",s", ",", .) %>%
-    # add spaces back in so more readable
-    gsub(",", ", ", .)
-
-  # add in as column
-  stab1$STROBEItem <- x
-
-  colnames(stab1)[colnames(stab1) == "rule"] <- "Rule"
-  colnames(stab1)[colnames(stab1) == "STROBEItem"] <- "STROBE Item" 
-  # remove all unnecessary cols
-  stab1 <- stab1[, -grep("variable|response", colnames(stab1))]
-
-  stab1$Rule <- gsub("^Rule = |^Rule =", "", stab1$Rule)
-
-  return(stab1)
-
-}
-
-stab1 <- create_stab1()
-
-pander(stab1, caption = "Supplementary Table 1. Rules for resolving conflicts", justify = c("left", "center"))
+```{r, echo = F}
+pander(kappa_sub, caption = "Supplementary Table 1. Kappa statistics for reporting quality of subdivisions of all STROBE items")
 ```
 
-## Supplementary Table 2
-
-```{r STable2, echo=F}
-
-# create labels for strobe items at subdivision level
-
-sub_labels <- read.csv("../outputs/extraction_dict.csv", stringsAsFactors = F, encoding = "UTF-8") %>%
-  .[grep(strobe_pattern, .$variable), ] %>%
-  .[-grep("_ev$", .$variable),]
-
-# create stabe
-
-create_stab2 <- function(){
-  stab2 <- df[, which(colnames(df) %in% c(strobe_cols_names))]
-  # convert all to characters so can easily reset values and table will just treat as nominal
-  for(i in colnames(stab2)){
-    stab2[[i]] <- as.character(stab2[[i]])
-  }
-
-  # set missing to na as won't make a difference to absolute numbers
-  stab2[is.na(stab2)|stab2 == "na"] <- "Not Applicable"
-  stab2[stab2 == "PartiallyExternal"] <- "Yes"
-
-  # pivot so can table frequencies
-  stab2 <- tidyr::pivot_longer(stab2, colnames(stab2), names_to = "strobe_item", values_to = "value") %>%
-    # reorder so follows strobe order
-    .[mixedorder(as.character(.$strobe_item)), ]
-
-  # remove s from strobe item
-  stab2$strobe_item <- gsub("^s|starred", "", stab2$strobe_item)
-
-  strobe_levels <-  unique(stab2$strobe_item)
-
-  # set levels as mixed order so order preserved in ggplot and pander
-  stab2$strobe_item<- factor(stab2$strobe_item, levels =  strobe_levels[gtools::mixedorder(strobe_levels)])
-
-  # order levels 
-  stab2$value <- as.factor(stab2$value)
-  levels(stab2$value) <- list(NotApplicable = "Not Applicable", No = "No", Partially = "Partially", Yes = "Yes")
-
-  return(stab2)
-}
-
-stab2 <- create_stab2()
-stab2_tab <- table(stab2$strobe_item, stab2$value, useNA = "ifany")
-
-png("Rplot10.png", width = 1300, height = 1500)
-
-# strobe item on x axis and plot % yes
-stab2 %>%ggplot(aes(x=strobe_item, y=..count.., fill = value)) + 
-  # blue fill
-  geom_bar() +
-  theme_classic() +
-  theme(text = element_text(size=30))+
-  #scale_x_discrete(labels= labels$x_labels) +
-  #scale_y_continuous(breaks = seq(0, nrow(df), by = 10)) + 
-  coord_flip() +
-  xlab("STROBE Item") +
-  ylab("Count") +
-  scale_fill_brewer(palette = "Greys") +
-  #geom_text(aes(label = count), vjust = 0, size = 8) +
-  theme(plot.title = element_text(size = 22)) 
-dev.off()
-
-pander(stab2_tab, caption = "Supplementary Table 2. Descriptive statistics for reporting quality of subdivisions of all STROBE items")
-```
-
-## Supplementary Table 3
-
-```{r, echo = F}
-pander(kappa_sub, caption = "Supplementary Table 3. Kappa statistics for reporting quality of subdivisions of all STROBE items")
-```
-
 # References
